---
author: "Gabriele Tomè, Toma Tebaldi"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmdformats::readthedown:
    css: custom_theme.css
    df_print: paged
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
params:
  doc_title: "Report of integrative pipeline for proteomics analysis"
  description: NA
  readPD_files: NA
  readMQ_files: NA
  file_input: NA
  file_prot: NA
  file_pep: NA
  signal_thr: "inf"
  fc_thr: "0.75"
  pval_thr: "0.05"
  batch_corr_exe: NA
  contr_design: NA
  prot_boxplot: NA
  run_enrich: NA
  run_STRING: NA
  pval_enrich_thr: "0.05"
  overlap_size_enrich_thr: 5
  enrich_filter_term: NA
  enrich_filter_DBs: NA
  dirOutput: NA
---

```{r settings, message=F, warning=F, echo=F,results='hide'}
# require("knitr")
# library("rstudioapi")
# library(stringr)
# #find and set the home directory depending on the position of the current Rmd
# PATH_R_FILE<-paste(str_split(getSourceEditorContext()$path, "/")[[1]][1:(length(str_split(getSourceEditorContext()$path, "/")[[1]])-2)],collapse = "/")
# opts_knit$set(root.dir = PATH_R_FILE,
#               dev="cairo.pdf")
# setwd(PATH_R_FILE)
# 
# #Create the output directory
# dirOutput_2="../outputResults/"
# dir.create(file.path("./",dirOutput_2), showWarnings = FALSE)
# currentTime = gsub(".*?([0-9]+).*?", "\\1", Sys.time())
# dirOutput_1=paste(currentTime,"/",sep = "")
# dir.create(file.path(dirOutput_2, dirOutput_1), showWarnings = FALSE)
# dirOutput=paste(dirOutput_2,dirOutput_1,sep = "")
dirOutput = params$dirOutput
#Auto-install needed library
list.of.packages <- c("ggrepel","tidyr","ggraph","graphlayouts","RColorBrewer",
                      "scales","igraph","data.table","qpdf","devtools",
                      "corrplot", "data.table", "ggplot2", "ggfortify","lazyeval",
                      "lubridate", "pheatmap", "reshape2","readr", "rlang",
                      "tibble", "dplyr", "tidyr", "wesanderson","WGCNA")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) suppressMessages(suppressWarnings({install.packages(new.packages, dependencies = T)}))
list.of.packages.bio <- c("biomaRt","PhosR","DEqMS","STRINGdb",
                          "GO.db", "impute", "preprocessCore", "pvca","sva")
new.packages <- list.of.packages.bio[!(list.of.packages.bio %in% installed.packages()[,"Package"])]
if(length(new.packages)) {
  if (!require("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  suppressMessages(suppressWarnings({BiocManager::install(new.packages)}))
}
list.of.packages <- c("geomnet","proBatch")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) {
  suppressMessages(suppressWarnings({devtools::install_github("sctyner/geomnet", dependencies = T)}))
  suppressMessages(suppressWarnings({devtools::install_github("symbioticMe/proBatch", dependencies = T)}))
}

#Load functions
suppressMessages(suppressWarnings({
  source("functions_2021.R") # new functions 2021
  #Library needed
  library('biomaRt')
  library('PhosR')
  library('DEqMS')
  library(ggrepel) # graphics
  library(tidyr)
  library(ggraph)
  library(graphlayouts)
  library(RColorBrewer) # Color palettes 
  library(scales) # heatmap colors scales
  library(geomnet) # network plots
  library(igraph) # Network analysis
  library(data.table)
  library(qpdf)
  library(STRINGdb)
  library("proBatch")
}))
  
options(dplyr.summarise.inform = FALSE)

#Set font and size
loadfonts(quiet=T)
bs<-11      # point sizes in ggplot
bf<-"Arial" # font family in ggplot

# Color definition
color_base<-c("dodgerblue2", "#E31A1C", # red
              "green4", "#6A3D9A", # purple
              "#FF7F00", # orange
              "black", "gold1", "skyblue2", "#FB9A99", # lt pink
              "palegreen2","#CAB2D6", #lt purple
              "#FDBF6F", #lt orange
              "#9af7c9","#c0a0e6","#f5939a","#5fefee","#e8d388","#8abdff","#e3bf7a","#48c0f6","#ccb86f","#fafeaf",
              "#04bfe4","#ea92a8","#25cfe9","#ffbb97","#62e8ff","#ffd29b","#8fcbff","#aec87f","#b2c6ff","#aedf9b",
              "#ffb3c8","#81b97c","#ffcef0","#54bcae","#ffcab3","#7dd7ff","#ffe4ad","#81b0da","#ebffc5","#a7a8d5",
              "#bdffd0","#d6d2ff","#9fb282","#a5d8ff","#c9a395","#9bfff7","#ffcbce","#70b6ca","#faffdd","#88b3bc",
              "#d6ffdf","#a3e5ff","#b1ac8f","#ccffff","#b4bb9f","#7eb7a5","#c5dbc3","#a8b9ff","#74dfbc","#f2ace7")
colour_vec<-color_base

all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(time_it = TRUE)
```

```{r load_settings, warning=FALSE, echo=F, message=FALSE, results='hide'}
message("Reading options")
#Load the excel file with  the basic settings
# input_settings <- tryCatch({suppressMessages({readxl::read_excel("options.xlsx", col_names = F)})},
#                            error=function(cond){stop("Error: \'options.xlsx\' not present in the current folder. Follow the instruction in the README.")})
# input_settings <- input_settings[rowSums(is.na(input_settings)) != ncol(input_settings), ]
# #Read title and description
# bindingIsLocked("params", env = .GlobalEnv)
# unlockBinding("params", env = .GlobalEnv)
# if(!is.na(input_settings$...2[1])){params$doc_title<-input_settings$...2[1]}
description<-params$description
# input_settings<-input_settings[3:nrow(input_settings),]
#Read Platform of previuos quantification
# sw_analysis <-input_settings$...2[1]
# readPD_files <- readMQ_files <- FALSE
# if(sw_analysis == "PD"){ readPD_files <- TRUE 
# }else if(sw_analysis == "MQ"){ readMQ_files <- TRUE
# }else { stop("Error: Incorrect plaftorm of quantification inserted. Write \'PD\' for Protein Discovery or \'MQ\' for MaxQuant.")}
readPD_files <- params$readPD_files
readMQ_files <- params$readMQ_files
#Read input directory
# dir_file_input<-input_settings$...2[2]
filenames<-list(params$file_input, params$file_pep, params$file_prot)
#Read input parameters for Limma
signal_thr=suppressWarnings(parse_double(params$signal_thr)) 
if(is.na(signal_thr)){signal_thr<- -Inf}
fc_thr=suppressWarnings(parse_double(params$fc_thr))
if(is.na(fc_thr)){
  message("Warning: parse error for Fold Change threshold. Set default paramenter to 0.75\n")
  fc_thr <- 0.75
}
pval_thr=suppressWarnings(parse_double(params$pval_thr))
if(is.na(pval_thr)){
  message("Warning: parse error for P.Value threshold. Set default paramenter to 5%\n")
  pval_thr <- 0.05
}
batch_corr_exe<-as.logical(params$batch_corr_exe)

# save(signal_thr, fc_thr, pval_thr, batch_corr_exe, readMQ_files, readPD_files, file="tmp.RData")
message(params$contr_design)
#Read the design
input_settings<-read_excel_allsheets(params$contr_design)[[1]]
contr_design<-data.frame(t(input_settings[1:nrow(input_settings),]))
contr_design<-data.frame(contr_design[,colSums(is.na(contr_design)) != nrow(contr_design) ])
if(ncol(contr_design) > 0){
  if(is.na(contr_design[3,])){
    contr_design[3,]<-color_base[1:ncol(contr_design)]
  }
  contr_design<-data.frame(t(fill(contr_design, names(contr_design), .direction = "down")))
  contr_design[,2]<-unlist(lapply(contr_design[,2], function(x){ if("-" %in% x){str_replace_all(x,"-","_")}else{x}}))
  formule_contrast<-(contr_design[,1])
  names(formule_contrast)<-contr_design[,2]
  color_constrast<-contr_design[,3]
  names(color_constrast)<-contr_design[,2]
  
  darker.col = function(color, how.much = 50){
    colorRampPalette(c(color, "black"))(100)[how.much]
  }
  color_constrast_dark<-c()
  for(col in names(color_constrast)){
    color_constrast_dark[col]<-darker.col(color_constrast[col])
  }
  color_constrast<-cbind(color_constrast,color_constrast_dark)
}else{
  stop("Error: At least one contrast design is required for the analysis.")
}

#Read Proteins List for control boxplot
prot_boxplot<-na.omit(strsplit(params$prot_boxplot, split = "\\;|\\,|\\ |\\; |\\, ")[[1]])
boxplot_prot_exe <- FALSE
if(length(prot_boxplot)>0){ boxplot_prot_exe<-TRUE }
#Read enrich parameter
doNextChunk_STRING <- doNextChunk <- run_enrich <- run_STRING <- FALSE
run_enrich <- as.logical(params$run_enrich)
run_STRING <- as.logical(params$run_STRING)
save(params, file = "tmp.RData")
pval_enrich_thr <- if(is.null(params$pval_enrich_thr)){0.05
                   }else if(is.double(params$pval_enrich_thr)){params$pval_enrich_thr
                   }else{parse_double(params$pval_enrich_thr)}
if(is.na(pval_enrich_thr)){
  message("Warning: parse error for p-value enrichment threshold. Set default paramenter to 0.05\n")
  pval_enrich_thr <- 0.05
}
overlap_size_enrich_thr=if(is.null(params$overlap_size_enrich_thr)){5
                         }else if(is.integer(params$overlap_size_enrich_thr)){params$overlap_size_enrich_thr
                         }else{parse_double(params$overlap_size_enrich_thr)}
if(is.na(overlap_size_enrich_thr)){
  message("Warning: parse error for Overlap size enrichment threshold. Set default paramenter to 5\n")
  overlap_size_enrich_thr <- 5
}

enrich_filter_term<-if(is.null(params$enrich_filter_term)){list()
}else if(params$enrich_filter_term == ""){list()
}else{na.omit(strsplit(params$enrich_filter_term, split = "\\;|\\,|\\ |\\; |\\, ")[[1]])}
enrich_filter_DBs<-if(is.null(params$enrich_filter_DBs)){list()
}else if(params$enrich_filter_DBs == ""){list()
}else{na.omit(params$enrich_filter_DBs)}

remove(input_settings,contr_design,color_constrast_dark)
```

```{r load_settings_old, warning=FALSE, echo=F, message=FALSE, results='hide', eval=FALSE}
message("Reading options")
#Load the excel file with  the basic settings
input_settings <- tryCatch({suppressMessages({readxl::read_excel("options.xlsx", col_names = F)})},
                           error=function(cond){stop("Error: \'options.xlsx\' not present in the current folder. Follow the instruction in the README.")})
input_settings <- input_settings[rowSums(is.na(input_settings)) != ncol(input_settings), ]
#Read title and description
bindingIsLocked("params", env = .GlobalEnv)
unlockBinding("params", env = .GlobalEnv)
if(!is.na(input_settings$...2[1])){params$doc_title<-input_settings$...2[1]}
description<-input_settings$...2[2]
input_settings<-input_settings[3:nrow(input_settings),]
#Read Platform of previuos quantification
sw_analysis <-input_settings$...2[1]
readPD_files <- readMQ_files <- FALSE
if(sw_analysis == "PD"){ readPD_files <- TRUE 
}else if(sw_analysis == "MQ"){ readMQ_files <- TRUE
}else { stop("Error: Incorrect plaftorm of quantification inserted. Write \'PD\' for Protein Discovery or \'MQ\' for MaxQuant.")}
#Read input directory
dir_file_input<-input_settings$...2[2]
#Read input parameters for Limma
signal_thr=suppressWarnings(parse_double(input_settings$...2[4])) 
if(is.na(signal_thr)){signal_thr<- -Inf}
fc_thr=suppressWarnings(parse_double(input_settings$...2[5]))
if(is.na(fc_thr)){
  message("Warning: parse error for Fold Change threshold. Set default paramenter to 0.75\n")
  fc_thr <- 0.75
}
pval_thr=suppressWarnings(parse_double(input_settings$...2[6]))
if(is.na(pval_thr)){
  message("Warning: parse error for P.Value threshold. Set default paramenter to 5%\n")
  pval_thr <- 0.05
}
batch_corr_exe<-as.logical(input_settings$...2[7])

#Read the design
contr_design<-data.frame(t(input_settings[10:nrow(input_settings),c(1,2,3)]))
contr_design<-data.frame(contr_design[,colSums(is.na(contr_design)) != nrow(contr_design) ])
if(ncol(contr_design) > 0){
  if(is.na(contr_design[3,])){
    contr_design[3,]<-color_base[1:ncol(contr_design)]
  }
  contr_design<-data.frame(t(fill(contr_design, names(contr_design), .direction = "down")))
  contr_design[,2]<-unlist(lapply(contr_design[,2], function(x){ if("-" %in% x){str_replace_all(x,"-","_")}else{x}}))
  formule_contrast<-(contr_design[,1])
  names(formule_contrast)<-contr_design[,2]
  color_constrast<-contr_design[,3]
  names(color_constrast)<-contr_design[,2]
  
  darker.col = function(color, how.much = 50){
    colorRampPalette(c(color, "black"))(100)[how.much]
  }
  color_constrast_dark<-c()
  for(col in names(color_constrast)){
    color_constrast_dark[col]<-darker.col(color_constrast[col])
  }
  color_constrast<-cbind(color_constrast,color_constrast_dark)
}else{
  stop("Error: At least one contrast design is required for the analysis.")
}

#Read Proteins List for control boxplot
prot_boxplot<-na.omit((input_settings[9:nrow(input_settings),c(5)])[[1]])
boxplot_prot_exe <- FALSE
if(length(prot_boxplot)>0){ boxplot_prot_exe<-TRUE }

#Read enrich parameter
doNextChunk_STRING <- doNextChunk <- run_enrich <- run_STRING <- FALSE
run_enrich <- as.logical(input_settings$...8[2])
run_STRING <- as.logical(input_settings$...8[3])

pval_enrich_thr <- suppressWarnings(parse_double(input_settings$...8[5]))
if(is.na(pval_enrich_thr)){
  message("Warning: parse error for p-value enrichment threshold. Set default paramenter to 0.05\n")
  pval_enrich_thr <- 0.05
}
overlap_size_enrich_thr=suppressWarnings(parse_double(input_settings$...8[6]))
if(is.na(overlap_size_enrich_thr)){
  message("Warning: parse error for Overlap size enrichment threshold. Set default paramenter to 5\n")
  overlap_size_enrich_thr <- 5
}

enrich_filter_term<-na.omit((input_settings[10:nrow(input_settings),c(7)])[[1]])
enrich_filter_DBs<-na.omit((input_settings[10:nrow(input_settings),c(8)])[[1]])

remove(input_settings,contr_design,color_constrast_dark)
```


---
title: `r params$doc_title`
---

`r description`

This report summarize the results obtained during the execution of the integrative pipeline for proteomics analysis. More results can be found in the folder.

## Details on the output files

- ```report.html```: complete report of the analysis with all pics and results of the enrichment.
- ```database_env_R.RData```: RData containing the essential variables for further analysis.
- ```normalised_intensity_table.xlsx```: excel file with the matrix used in the analysis. In sheet *Normalized proteins matrix intensity* is written the normalized and imputed version of the original intensities of each protein in each sample. In sheet *Normalized peptides matrix intensity* is written the normalized and imputed version of the original intensities of each peptide in each sample. In sheet *Normalized proteins intensity group by* is written the average and the standard deviation of each condition described in the Input file for each protein. In sheet *Normalized peptides intensity group by* is written the average and the standard deviation of each condition described in the Input file for each peptide.
- ```DEPs_table.xlsx```: summary of the result of Limma. In sheet *DEPs_proteins*, Column *id* is the name of the protein. In sheet *DEPs_peptides*, column *Accession* is the UniprotID of protein, column *Description* is the description of protein, *GeneName* is Gene Symbol, *Annotated Sequence* is the sequence of peptides, *Modifications* is the modification that affect the peptides, *Position in Master Proteins* is the UniprotID with the position on the sequence. For each contrast rule are present 5 columns:
  - **class**: is "+" if the protein is up-regulated following the parameters set in **options.xlsx**, "-" if the protein is down-regulated, "=" otherwise
  - **log2_FC**: is the log2 Fold Change of the contrast
  - **p_val**: is the p. value
  - **p_adj**: is the adjusted p. value
  - **log2_expr**: is the signal log2 expression
  
- ```general_mds_proteins.pdf```: MDS of the samples by proteins.
- ```general_mds_peptides.pdf```: MDS of the samples by peptides.

```{r print_descr_batch, warning=FALSE, echo=F, message=FALSE, eval=batch_corr_exe, results='asis'}
cat("- ```corrected_mds_proteins.pdf```: MDS of the samples by proteins after the bacth effect correction \n")
cat("- ```corrected_mds_peptides.pdf```: MDS of the samples by peptides after the bacth effect correction \n")
```

```{r print_descr_control, warning=FALSE, echo=F, message=FALSE, eval=prot_boxplot, results='asis'}
cat("- ```boxplot_control_intensity.pdf```: boxplot for the control of the intensities of selected proteins \n")
```

- ```DEPs_proteins_count_barplot.pdf```: number of DEPs find for each contrast design based on proteins. Light grey are the down-regulated, dark grey are the up-regulated genes.
- ```DEPs_peptides_count_barplot.pdf```: number of DEPs find for each contrast design based on peptides. Light grey are the down-regulated, dark grey are the up-regulated genes.
- ```DEPs_proteins_mds.pdf```: MDS of the DEPs discovered based on proteins.
- ```DEPs_peptides_mds.pdf```: MDS of the DEPs discovered based on peptides.

```{r print_descr_enrich, warning=FALSE, echo=F, message=FALSE, eval=run_enrich, results='asis'}
cat("- ```DEPs_enrichment_data.RData```: result of the enrichment made with EnrichR \n")
cat("- ```DEPs_enrichment_table.xlsx```: excel with only the enriched terms with an high significance (P.Value < 0.05, Overlap Size >= 4) \n")
cat("- ```enrich``` directory: plots of the enrichment data under the parameters described in the **option.xlsx** file \n")  
cat("  - **DBs_enr.pdf**: dot matrix of the enrichment result divided in up- and down-regulated genes. Filtered by the datasets choose in the **options** file. \n")
cat("  - **terms_enr.pdf**: dot matrix of the enrichment result divided in up- and down-regulated genes. Filtered by the terms written in the **options** file. \n")
cat("  - **DBs_enr_all.pdf**: dot matrix of the enrichment result combining the up- and down- regulated genes in a single list. Filtered by the datasets choose in the **options** file. \n")
cat("  - **terms_enr_all.pdf**: dot matrix of the enrichment result combining the up- and down- regulated genes in a single list. Filtered by the terms written in the **options** file. \n")
```

```{r print_descr_STRING, warning=FALSE, echo=F, message=FALSE, eval=run_STRING, results='asis'}
cat("- ```network``` directory: plots of the up- and down- regulated genes based on STRINGdb. For each contrast design are create 2 files: \n")
#  <!-- - **Prefilter**: histogram representing the number of genes in each community discovered. -->
cat("  - **Postfilter**: histogram representing the number of genes in each community discovered after a filter on the genes and on the links. \n")
cat("  - **Graph**: visual representation of the network discovered in two layouts: *fr* and *kk*. \n")
```


## Filter thresholds

```{r print_settings, warning=FALSE, echo=F, message=FALSE, results='asis'}
cat(" \n \n")
#Print summary of filter and contrast design
cat("**Threshold used during the analysis:** ")
cat(" \n \n")
cat(paste0("- Log2 Fold Change (FC) threshold: FC > ",fc_thr,", FC < -",fc_thr," "))
cat(" \n")
cat(paste0("- P-Value threshold: p.val < ",pval_thr," "))
cat(" \n")
if(signal_thr != -Inf){cat(paste0("- Signal log2 intesity threshold: signal > ",signal_thr," "))}
cat(" \n")
if(run_enrich){cat(paste0("- P-Value threshold for enrichment analysis: p.val < ",pval_enrich_thr," "))}
cat(" \n")
if(run_enrich){cat(paste0("- Overlap size threshold for enrichment analysis: OverlapSize > ",overlap_size_enrich_thr," "))}
cat(" \n")
if(batch_corr_exe){cat("- DONE batch effect correction.")}
```

## Constrast design used for the differential analysis
The following formulas are the comparison on which is performed the differential analysis.

```{r print_settings_design, warning=FALSE, echo=F, message=FALSE, results='asis'}
cat("**Contrast design formula are:** ")
cat(" \n")
for (i in 1:length(formule_contrast)){
  cat(paste0("- ",names(formule_contrast[i]),": ",formule_contrast[i]," "))
  cat(" \n")
}
cat(" \n")

```

## 1. Annotation sample from input file
The table summarize the input samples. Each case has: 

- a sample name in column "sample", 
- the associated condition of study in column "condition",
- the color of that sample in column "color".


```{r load_data_from_PD, eval=readPD_files,echo=F, message=F, warning=F, results='asis'}
message("Reading files from MS...")
#Read all files in the input folder
template_files=c("Input","pep","prot")
filenames <- list("Input"=params$file_input,"prot"=params$file_prot,"pep"=params$file_pep)
input_files<-list("FILES")
for (f in template_files){
  input_files[f] <- tryCatch({
    read_excel_allsheets(filenames[[1]])
    # read_excel_allsheets(paste0(dir_file_input,filenames[which(grepl(f,filenames) & !grepl('~', filenames))]))
  }, error=function(cond){
    stop(paste0("Error: missing file. The file \'",f,"\' is missing or not have the pattern in the filename or there are duplicates files."))
  })
}
names(input_files)<-c("dir","Input","PEP","PROT")
message("Data acquired")


#Clean files and merge
colToKeep<-intersect(colnames(input_files$Input), c("File ID","File Name","Condition","Color","Sample","MS_batch"))
input_files$Input <- input_files$Input[,colToKeep]
if(!("Condition" %in% colnames(input_files$Input))){ stop("Error: \'Condition\' column missin in \'Input\' file.") }
if(batch_corr_exe & !("MS_batch" %in% colnames(input_files$Input))){ stop("Error: \'MS_batch\' column missin in \'Input\' file with batch correction activated") }
input_files$Input$`File Name`<- lapply(input_files$Input$`File Name`,
                                                   function(x){tail(unlist(strsplit(x,"\\",fixed=T)),n=1)})

input_files$Input$`File Name`<-mf(unlist(input_files$Input$`File Name`),"_")
#Prepare PROT table with only column Accession and Dscription, extract fro Description the GeneName and remove contaminants
if(is_empty(which(grepl("Keratin|keratin",input_files$PROT$Description) | 
                  grepl("CON_",input_files$PROT$Accession)))){
  input_files$PROT <- na.omit(input_files$PROT[,c("Accession","Description")])
}else{
  input_files$PROT <- na.omit(input_files$PROT[-which(grepl("Keratin|keratin",input_files$PROT$Description) |
                                              grepl("CON_",input_files$PROT$Accession))
                                       ,c("Accession","Description")])
}
input_files$PROT$GeneName<-lapply(input_files$PROT$Description, function(x){gsub(".*GN=(.+) PE=.*", "\\1", x)}) %>% unlist()
taxonomy_NCBI<-lapply(input_files$PROT$Description, function(x){gsub(".*OX=(.+) GN=.*", "\\1", x)}) %>% unlist()
taxonomy_NCBI<-strtoi(names(sort(table(taxonomy_NCBI), decreasing = T))[1])
#Mantain only the first UNIPROT code
input_files$PEP$`Master Protein Accessions`<-lapply(input_files$PEP$`Master Protein Accessions`,
                                                                function(x){unlist(str_split(x, ";"))[1]}) %>% unlist()
#Verify if the Abundance start in 5 column in PEP table, if not remove extra column
if(!grepl("Abundance",colnames(input_files$PEP)[5])){
  input_files$PEP<-input_files$PEP[,c("Annotated Sequence","Modifications",
                                      "Master Protein Accessions","Positions in Master Proteins",
                                      colnames(input_files$PEP)[which(grepl("Abundance",colnames(input_files$PEP)))])]
}

#merge by uniprot the prot description and the pep table
uniprot_to_take<-merge(input_files$PROT, input_files$PEP, by.x = "Accession", by.y = "Master Protein Accessions")
input_files$PD_PEP_matrix<-merge(input_files$PROT[which(input_files$PROT$Accession %in% uniprot_to_take$Accession),],
                                 input_files$PEP, by.x = "Accession", by.y = "Master Protein Accessions")
input_files$PD_PEP_matrix<-input_files$PD_PEP_matrix[,c(c("Accession","Description","GeneName","Annotated Sequence",
                                                          "Modifications","Positions in Master Proteins"),
                                                        lapply(input_files$Input$`File ID`, function(x){
                                                          colnames(input_files$PD_PEP_matrix)[grepl(x, colnames(input_files$PD_PEP_matrix))]
                                                          }) %>% unlist())]
if("TRUE" %in% grepl("\\.", colnames(input_files$PD_PEP_matrix))){
  input_files$PD_PEP_matrix<-input_files$PD_PEP_matrix[,-which(colnames(input_files$PD_PEP_matrix) %in%
                                                               c(colnames(input_files$PD_PEP_matrix)[grepl("\\.", 
                                                                                            colnames(input_files$PD_PEP_matrix))]))]
}
colnames(input_files$PD_PEP_matrix)<-c(c("Accession","Description","GeneName","Annotated Sequence","Modifications","Position in Master Proteins"),
                                        unlist(lapply(colnames(input_files$PD_PEP_matrix), function(x){
                                           lapply(input_files$Input$`File ID`, function(y){if(str_detect(x,paste0(y,":"))){
                                               input_files$Input$`File Name`[which(input_files$Input$`File ID` %in% y)]
                                           }})
                                       })))
# input_files$PD_PEP_matrix$GeneName<-str_split_fixed(str_split_fixed(input_files$PD_PEP_matrix$GeneName,"-",n=2)[,1],"[.]",n=2)[,1] %>% toupper()
input_files$PD_PEP_matrix$GeneName<-str_split_fixed(str_split_fixed(input_files$PD_PEP_matrix$GeneName,"-",n=2)[,1],"[.]",n=2)[,1]
#Maintain only the new complete matrix of intensity
psm_sig_raw<-input_files$PD_PEP_matrix
c_anno<-input_files$Input
if(!("Sample" %in% colnames(c_anno))){ 
  colnames(c_anno)[2]<-"sample" 
}else{
  colnames(psm_sig_raw)<-c(c("Accession","Description","GeneName","Annotated Sequence","Modifications","Position in Master Proteins"),
                           c_anno$Sample)
}
colnames(c_anno)<-tolower(colnames(c_anno))

rownames(c_anno)<-c_anno$sample
if("color" %in% colnames(c_anno)){ 
  colour_vec<-na.omit(c_anno[colnames(psm_sig_raw),]$color)
  names(colour_vec)<-na.omit(c_anno[colnames(psm_sig_raw),]$sample)
}

remove(uniprot_to_take,input_files)
#Print Annotation table
kable(c_anno)

```

```{r load_data_from_MQ, eval=readMQ_files,echo=F, message=F, warning=F, results='asis'}
message("Reading files from MS...")
#Read all files in the input folder
template_files=c("Input","pep","prot")
filenames <- list("Input"=params$file_input,"prot"=params$file_prot,"pep"=params$file_pep)
input_files<-list("FILES"="in_files")
# save(template_files, filenames, input_files, file = "tmp.RData")
for (f in template_files){
  if(f == template_files[1]){
    input_files[f] <- tryCatch({
      read_excel_allsheets(filenames[f][[1]])
      # read_excel_allsheets(paste0(dir_file_input,filenames[which(grepl(f,filenames) & !grepl('~', filenames))]))
    }, error=function(cond){
      stop(paste0("Error: missing file. The file \'",f,"\' is missing or not have the pattern in the filename or there are duplicates files."))
    })
  }else{
    input_files[f] <- tryCatch({
      list(read_tsv(filenames[f][[1]]))

      # list(read_tsv(paste0(dir_file_input,filenames[which(grepl(f,filenames) & !grepl('~', filenames))])))
    }, error=function(cond){
      stop(paste0("Error: missing file. The file \'",f,"\' is missing or not have the pattern in the filename or there are duplicates files."))
    })
  }
}
names(input_files)<-c("dir","Input","PEP","PROT")
message("Data acquired")


#Clean files and merge
colToKeep<-intersect(colnames(input_files$Input), c("Condition","Color","Sample"))
input_files$Input <- input_files$Input[,colToKeep]
if(!("Condition" %in% colnames(input_files$Input))){ stop("Error: \'Condition\' column missin in \'Input\' file.") }
if(!("Sample" %in% colnames(input_files$Input))){ stop("Error: \'Sample\' column missin in \'Input\' file.") }
#Prepare PROT table with only column Accession and Dscription, extract fro Description the GeneName and remove contaminants
input_files$PROT <- input_files$PROT[-which(grepl("Keratin|keratin",input_files$PROT$`Fasta headers`) |
                                            grepl("CON_",input_files$PROT$`Majority protein IDs`))
                                     ,c("Majority protein IDs","Fasta headers")]
input_files$PROT$GeneName<-lapply(input_files$PROT$`Fasta headers`, function(x){gsub(".*GN=(.+) PE=.*", "\\1", x)}) %>% unlist()
input_files$PROT$GeneName[which(str_detect(input_files$PROT$GeneName,";"))] <- lapply(
                                                                        input_files$PROT$GeneName[which(str_detect(input_files$PROT$GeneName,";"))], 
                                                                                  function(x){unlist(str_split(x, ";"))[1]}) %>% unlist()

taxonomy_NCBI<-lapply(input_files$PROT$`Fasta headers`, function(x){gsub(".*OX=(.+) GN=.*", "\\1", x)}) %>% unlist()
taxonomy_NCBI<-strtoi(names(sort(table(taxonomy_NCBI), decreasing = T))[1])
if(is.na(taxonomy_NCBI)){
  taxonomy_NCBI<-9606
}
#Mantain only the first UNIPROT code
# input_files$PEP$`Master Protein Accessions`<-lapply(input_files$PEP$`Master Protein Accessions`,
#                                                                 function(x){unlist(str_split(x, ";"))[1]}) %>% unlist()
input_files$PEP<-input_files$PEP[,c("Proteins","Sequence","Start position","End position",
                                    colnames(input_files$PEP)[which(grepl("Intensity ",colnames(input_files$PEP)))])]
input_files$PEP$`End position` <- str_c(input_files$PROT$GeneName,input_files$PEP$`Start position`, input_files$PEP$`End position`, sep = "_")
input_files$PEP$`Start position` <- c(NA)

#merge by uniprot the prot description and the pep table
input_files$PD_PEP_matrix<-merge(input_files$PROT, input_files$PEP, by.x = "Majority protein IDs", by.y = "Proteins")
colnames(input_files$PD_PEP_matrix)<-c(c("Accession","Description","GeneName",
                                         "Annotated Sequence","Modifications","Position in Master Proteins"),
                                        unlist(lapply(colnames(input_files$PD_PEP_matrix), function(x){ if(str_detect(x,"Intensity")){
                                               str_split(x," ")[[1]][2]
                                           }})
                                       ))
# input_files$PD_PEP_matrix$GeneName<-str_split_fixed(str_split_fixed(input_files$PD_PEP_matrix$GeneName,"-",n=2)[,1],"[.]",n=2)[,1] %>% toupper()
input_files$PD_PEP_matrix$GeneName<-str_split_fixed(str_split_fixed(input_files$PD_PEP_matrix$GeneName,"-",n=2)[,1],"[.]",n=2)[,1]
#Maintain only the new complete matrix of intensity
psm_sig_raw<-input_files$PD_PEP_matrix
c_anno<-input_files$Input
colnames(c_anno)<-tolower(colnames(c_anno))

rownames(c_anno)<-c_anno$sample
if("color" %in% colnames(c_anno)){ 
  colour_vec<-na.omit(c_anno[colnames(psm_sig_raw),]$color)
  names(colour_vec)<-na.omit(c_anno[colnames(psm_sig_raw),]$sample)
}

remove(uniprot_to_take,input_files)
#Print Annotation table
kable(c_anno)

```

```{r peptide_annotation, message=F, warning=F, echo=F, eval=T}
psm_sig_raw$`Position in Master Proteins`<-gsub(" ", "_", psm_sig_raw$`Position in Master Proteins`)
psm_sig_raw$Modifications<-gsub(" ", "_", psm_sig_raw$Modifications)

if(readPD_files){
  #For peptides analysis
  # psm_anno_raw <- data.frame("accession"=psm_sig_raw$Accession,
  #                            "symbol"=psm_sig_raw$GeneName,
  #                            "sequence"=psm_sig_raw$`Annotated Sequence`,
  #                            "modifications"=psm_sig_raw$Modifications)
  #For protein groups analysis
  psm_anno_raw <- data.frame("symbol"=psm_sig_raw$GeneName,
                             "sequence"=psm_sig_raw$`Annotated Sequence`,
                             "modifications"=psm_sig_raw$Modifications)
  psm_anno_raw$id<-paste(psm_anno_raw$symbol,psm_anno_raw$sequence,psm_anno_raw$modifications,sep="_")
}else{
  psm_anno_raw <- data.frame("symbol"=psm_sig_raw$GeneName,
                             "sequence"=psm_sig_raw$`Annotated Sequence`,
                             "modifications"=psm_sig_raw$Modifications)
  psm_anno_raw$id<-paste(psm_anno_raw$symbol,psm_anno_raw$sequence,psm_anno_raw$modifications,sep="_")
}
rownames(psm_sig_raw)<-rownames(psm_anno_raw)<-paste(psm_anno_raw$symbol,seq(1:nrow(psm_anno_raw)),sep="_")
psm_anno_raw$old_id<-rownames(psm_anno_raw)

# create peptide names linked to symbols
psm_anno_raw$row <- seq(1,nrow(psm_anno_raw))
psm_anno_raw <- psm_anno_raw %>% dplyr::group_by(symbol) %>% dplyr::mutate("min"=min(row),"card"=n()) %>% ungroup()
psm_anno_raw$rank<- psm_anno_raw$row - psm_anno_raw$min +1
psm_anno_raw$id<- paste(psm_anno_raw$symbol,psm_anno_raw$rank,psm_anno_raw$card,sep="_")

psm_sig_raw<-psm_sig_raw[psm_anno_raw$old_id,]
rownames(psm_sig_raw)<-psm_anno_raw$id

psm_anno_raw<-as.data.frame(psm_anno_raw)
rownames(psm_anno_raw)<- psm_anno_raw$id

#Divide the table info from the matrix of peptide intensity
psm_peptide_table<-psm_sig_raw[,c(1:6)]
rownames(psm_peptide_table)<-paste(psm_peptide_table$`Position in Master Proteins`,
                                   psm_peptide_table$Modifications,
                                   psm_peptide_table$`Annotated Sequence`,sep = "_")
psm_sig_prot_raw<-psm_sig_raw[,c(7:ncol(psm_sig_raw))]
psm_sig_pet_raw<-psm_sig_raw[,c(7:ncol(psm_sig_raw))]
rownames(psm_sig_pet_raw)<-rownames(psm_peptide_table)

#For Peptide analysis
# rownames(psm_sig_raw)<-make.names(psm_anno_raw$accession,unique=T)
# rownames(psm_table_info)<-make.names(psm_table_info$Accession,unique=T)
# rownames(psm_anno_raw)<-make.names(psm_anno_raw$accession,unique=T)

```

## 2. Analysis

Raw data consist in the area under the peak (m/z, time, number of events) calculated for PSM (PSMs: peptide-to-spectrum matches).

-   PSMs not consistently present in all replicates of at least 1 condition will be removed.
-   Contaminants are removed.

The insentities are transformed in log2 scale.

```{r preprocess_sig_raw, echo=F, message=F, warning=F}
#Preprocess Protein intensities
psm_sig_prot_raw[psm_sig_prot_raw == 0] <- NA # transform 0s into NAs

sig_thr<-0 # NA threshold

psm_long_df <- psm_sig_prot_raw %>% mutate(id=as.factor(rownames(psm_sig_prot_raw)))
psm_long_df <- psm_long_df %>% pivot_longer(-id, names_to = "sample", values_to = "counts")
psm_long_df <- suppressMessages(left_join(psm_long_df,c_anno)) 

psm_filter_df <- psm_long_df %>% dplyr::group_by(id,condition, .drop=FALSE) %>% dplyr::summarise("min_c"=sum(is.na(counts))) %>% ungroup()

psm_filter_df <- psm_filter_df %>% dplyr::group_by(id,.drop=FALSE) %>% dplyr::summarise("passes_c"= sum(min_c <= sig_thr)) %>% ungroup()

filter_id <- as.character(psm_filter_df %>% dplyr::filter(passes_c > 0) %>% pull(id))

psm_sig_prot_df <-psm_sig_prot_raw[filter_id,]
psm_anno_df <- psm_anno_raw[filter_id,]
# psm_table_info <- psm_table_info[filter_id,]

#Log2
psm_log_prot_df <- psm_sig_prot_df
psm_log_prot_df <- psm_sig_prot_df %>% log2()

#Clean old variables
remove(psm_anno_raw,psm_sig_prot_raw,psm_long_df,psm_filter_df,filter_id,sig_thr)

#Preprocess peptide intensities
psm_sig_pet_raw[psm_sig_pet_raw == 0] <- NA # transform 0s into NAs
sig_thr<-0 # NA threshold

psm_long_df <- psm_sig_pet_raw %>% mutate(id=as.factor(rownames(psm_sig_pet_raw)))
psm_long_df <- psm_long_df %>% pivot_longer(-id, names_to = "sample", values_to = "counts")
psm_long_df <- suppressMessages(left_join(psm_long_df,c_anno)) 

psm_filter_df <- psm_long_df %>% dplyr::group_by(id,condition, .drop=FALSE) %>% dplyr::summarise("min_c"=sum(is.na(counts))) %>% ungroup()
psm_filter_df <- psm_filter_df %>% dplyr::group_by(id,.drop=FALSE) %>% dplyr::summarise("passes_c"= sum(min_c <= sig_thr)) %>% ungroup()
filter_id <- as.character(psm_filter_df %>% dplyr::filter(passes_c > 0) %>% pull(id))

psm_sig_pet_df <-psm_sig_pet_raw[filter_id,]
psm_peptide_table<-psm_peptide_table[filter_id,]

#Determine column tryptic
peptides_df <- psm_peptide_table[,c("Accession", "Annotated Sequence")]
peptides_df$preAA <- str_sub(str_extract(str_split_fixed(peptides_df$`Annotated Sequence`,regex("\\."), n = 3)[,1], regex("\\[\\w+\\]")), 
                             2, 
                             str_length(str_extract(str_split_fixed(peptides_df$`Annotated Sequence`,regex("\\."), 
                                                                    n = 3)[,1], regex("\\[\\w+\\]")))-1)

peptides_df$endAA <- str_sub(str_split_fixed(peptides_df$`Annotated Sequence`,regex("\\."), n = 3)[,2], -1, -1)
peptides_df$postAA <- str_sub(str_extract(str_split_fixed(peptides_df$`Annotated Sequence`,regex("\\."), n = 3)[,3], regex("\\[\\w+\\]")), 
                              2, 
                              str_length(str_extract(str_split_fixed(peptides_df$`Annotated Sequence`,regex("\\."), 
                                                                     n = 3)[,3], regex("\\[\\w+\\]")))-1)

peptides_df$fully_TRI = (peptides_df$preAA %in% c("K","R") &
                           peptides_df$endAA %in% c("K","R") &
                           (peptides_df$postAA != "P" | is.na(peptides_df$postAA)))
peptides_df$NSEMI_TRI = (peptides_df$preAA %in% c("K","R") &
                           !peptides_df$endAA %in% c("K","R") &
                           (peptides_df$postAA != "P" | is.na(peptides_df$postAA)))
peptides_df$CSEMI_TRI = (!peptides_df$preAA %in% c("K","R") &
                           peptides_df$endAA %in% c("K","R") &
                           (peptides_df$postAA != "P" | is.na(peptides_df$postAA)))
peptides_df$non_TRI = (!peptides_df$fully_TRI & !peptides_df$NSEMI_TRI & !peptides_df$CSEMI_TRI)

peptides_df$tryptic_cond[peptides_df$fully_TRI] <- "fully tryptic"
peptides_df$tryptic_cond[peptides_df$NSEMI_TRI] <- "N-semi tryptic"
peptides_df$tryptic_cond[peptides_df$CSEMI_TRI] <- "C-semi tryptic"
peptides_df$tryptic_cond[peptides_df$non_TRI] <- "non tryptic"

psm_peptide_table$tryptic_cond <- peptides_df$tryptic_cond

#Log2
psm_log_pet_df <- psm_sig_pet_df
psm_log_pet_df <- psm_sig_pet_df %>% log2()

remove(psm_sig_pet_raw,psm_sig_raw,psm_long_df,psm_filter_df,filter_id,sig_thr)
```

#### 2.1. Log2 signal distributions before imputation
The box plot represents the distribution of the intensity values for each samples in the raw data. This figure is generated before the normalization and imputation of the data.

```{r raw_distributions, message=F, warning=F, echo=F}

psm_long_df <- psm_log_prot_df %>% mutate(id=as.factor(rownames(psm_log_prot_df))) %>% pivot_longer(-id, names_to = "sample", values_to = "log2_intensity")
psm_long_df <- suppressMessages(left_join(psm_long_df,c_anno))


hs<-ggplot(data=psm_long_df, aes(x=sample,y=log2_intensity,fill=sample,colour=sample))+
  #geom_hline(yintercept = 0, lty=2)+
  coord_flip()+
  geom_violin(alpha=0.5,scale="width",trim = F,colour=NA)+
  geom_boxplot(alpha=1,fill="white",width=0.2,outlier.shape = NA, notch = T)+
  theme_bw()+
  theme(legend.position = "none",axis.title.y = element_blank()) +
  labs(x = "Log2 Intensity") +
  scale_fill_manual(values=(colour_vec)) +
  scale_colour_manual(values=(colour_vec)) +
  theme(panel.grid.minor = element_blank(),panel.grid.major.y = element_blank())
  
hs

```

#### 2.2. Percentage of numeric value in peptide intensity matrix for each sample
The next graph visualizes the percentage of intensities in the matrix. The percentage of intensity values are counted for each sample separately.

```{r numeric_numbers, echo=F, message=F, warning=F}

numeric_df<- c_anno
numeric_df$numeric_values<-colSums(is.na(psm_sig_prot_df))
numeric_df$numeric_percentage<- 100-(numeric_df$numeric_values/nrow(psm_sig_prot_df)*100)
setnames(numeric_df,"numeric_percentage","Percentage of known intensity (%)")
hs<-ggplot(data=numeric_df, aes(x=sample,y=`Percentage of known intensity (%)`,fill=sample,colour=sample))+
  #geom_hline(yintercept = 0, lty=2)+
  coord_flip()+
  geom_bar(stat="identity",width=.7,alpha=.8)+
  theme_bw()+
  theme(legend.position = "none",axis.title.y = element_blank())+
  labs(x = "Percentage of known intensity (%)") +
  scale_fill_manual(values=(colour_vec)) +
  scale_colour_manual(values=(colour_vec)) +
  theme(panel.grid.minor = element_blank(),panel.grid.major.y = element_blank())
  
hs

remove(hs,numeric_df)
```

To perform the double analysis at protein- and peptide-level, two normalization function are applied.
For the protein study, PSMs values were summarized into proteins and normalized across samples, with the "medianSweeping" function contained in the Bioconductor package [DEqMS](<https://www.bioconductor.org/packages/release/bioc/html/DEqMS.html>).
At the peptide level, the normalization is done by the function equalMedianNormalization from  [DEqMS](<https://www.bioconductor.org/packages/release/bioc/html/DEqMS.html>), which normalizes to have equal medians in all samples.

```{r prot_summarization, warning=F, message=F, echo=F}
message("Median normalization in progress...")
gene_anno_df <- psm_anno_df %>% dplyr::select(symbol) %>% unique()

dat_psm <- cbind(psm_anno_df[,c("id","symbol")],psm_log_prot_df)
#For peptides analysis
# dat_psm <- cbind(rownames(psm_anno_df),rownames(psm_anno_df),psm_log_df)

# 4 methods to summarise peptides into proteins: 
# medianSweeping (also correct intensities so that the median of all proteins is 0)
# medianSummary
# medpolishSummary
# farmsSummary

#Summarize into proteins and normalize by median
dat_gene = DEqMS::medianSweeping(dat_psm,group_col = 2)
remove(gene_anno_df)

# Median centering log2 table for peptides
dat_pep = DEqMS::equalMedianNormalization(psm_log_pet_df)

message("Normalization DONE.")
```

The principal method is based on the PhosR package. It is a package for comprehensive analysis of phosphoproteomic data. Its core components are data processing and downstream analysis.
PhosR is always applied but in some cases, it does not work, for example, if there is only 1 sample in a specific condition. In this rare case, it uses a simpler round imputation. It calculates the mean and the standard deviation of the matrix and it generates the imputed values based on the gaussian function derived.

```{r imputation, echo=F, message=F, warning=F, results='asis'}
message("Imputation of intensity matrix in progress...")
dat_gene<-tryCatch(
  {
    #Imputation with Phospho
    ppe <- PhosphoExperiment(assays = list(Quantification = as.matrix(dat_gene)))
    grps <- as.factor(c_anno$condition)
    ppe <- scImpute(ppe, 0.6, grps)
    ppe <- tImpute(ppe, assay = "imputed")
    dat_gene[is.na(dat_gene)]<-ppe@assays@data@listData$imputed[is.na(dat_gene)]
    cat("##### Done the *PhosR* imputation for proteins \n")
    dat_gene
  },
  error=function(cond){
    # Rounds of imputation
    mat_i <- dat_gene
    col_sele<-colnames(mat_i)
    
    vec_all<-unlist(mat_i[,col_sele]) # convert to vector
    vec_numeric<-vec_all[-which(is.na(vec_all))] # remove NAs 
    orig_stats<-c(mean(vec_numeric),sd(vec_numeric)) # calculate mean and sd
    imp_stats<-c(orig_stats[1]-(orig_stats[2]*1.8),orig_stats[2]*0.3) # imputed mean = mean-1.8*sd, imputed sd = sd*0.3
    
    set.seed(42)
    imp_matrix<-matrix(rnorm(length(vec_all), mean = imp_stats[1], sd = imp_stats[2]), ncol=ncol(mat_i)) # create matrix of imputed values
    dat_gene[is.na(dat_gene)]<-imp_matrix[is.na(dat_gene)]
    cat("##### Done *round* imputation for proteins \n")
    dat_gene
  }
)

remove(ppe,grps)

dat_pep<-tryCatch(
  {
    #Imputation with Phospho
    ppe <- PhosphoExperiment(assays = list(Quantification = as.matrix(dat_pep)))
    grps <- as.factor(c_anno$condition)
    ppe <- scImpute(ppe, 0.6, grps)
    ppe <- tImpute(ppe, assay = "imputed")
    dat_pep[is.na(dat_pep)]<-ppe@assays@data@listData$imputed[is.na(dat_pep)]
    cat("##### Done the *PhosR* imputation for peptides \n")
    dat_pep
  },
  error=function(cond){
    # Rounds of imputation
    mat_i <- dat_pep
    col_sele<-colnames(mat_i)
    
    vec_all<-unlist(mat_i[,col_sele]) # convert to vector
    vec_numeric<-vec_all[-which(is.na(vec_all))] # remove NAs 
    orig_stats<-c(mean(vec_numeric),sd(vec_numeric)) # calculate mean and sd
    imp_stats<-c(orig_stats[1]-(orig_stats[2]*1.8),orig_stats[2]*0.3) # imputed mean = mean-1.8*sd, imputed sd = sd*0.3
    
    set.seed(42)
    imp_matrix<-matrix(rnorm(length(vec_all), mean = imp_stats[1], sd = imp_stats[2]), ncol=ncol(mat_i)) # create matrix of imputed values
    dat_pep[is.na(dat_pep)]<-imp_matrix[is.na(dat_pep)]
    cat("##### Done *round* imputation for peptides \n")
    dat_pep
  }
)

remove(ppe,grps)


```


```{r pre_distributions, message=F, warning=F, echo=F, eval=F}

cat("#### 2.3. Distribution after the imputation and medianSweeping")

psm_long_df <- psm_log_prot_df %>% mutate(id=as.factor(rownames(psm_log_prot_df))) %>% pivot_longer(-id, names_to = "sample", values_to = "log2_intensity")
psm_long_df <- suppressMessages(left_join(psm_long_df,c_anno))


hs<-ggplot(data=psm_long_df, aes(x=sample,y=log2_intensity,fill=sample,colour=sample))+
  #geom_hline(yintercept = 0, lty=2)+
  coord_flip()+
  geom_violin(alpha=0.5,scale="width",trim = F,colour=NA)+
  geom_boxplot(alpha=1,fill="white",width=0.2,outlier.shape = NA, notch = T)+
  theme_bw()+
  theme(legend.position = "none",axis.title.y = element_blank())+
  scale_fill_manual(values=(colour_vec)) +
  scale_colour_manual(values=(colour_vec)) +
  theme(panel.grid.minor = element_blank(),panel.grid.major.y = element_blank())
  
hs

remove(hs,psm_long_df)

```

```{r sig_tables, message=F, warning=F, echo=F, results='hide'}
#Save proteins table
expr_mat<-dat_gene

expr_l_df <- expr_mat %>% mutate(id=rownames(expr_mat)) %>% pivot_longer(-id, names_to = "sample", values_to = "expr")
expr_l_df <- suppressMessages(left_join(expr_l_df,c_anno))

#Add Standard Deviation column and average
expr_cond_df<-expr_l_df %>% dplyr::group_by(condition,id) %>%
            dplyr::summarise(N=n(),avg=mean(expr),sd=sd(expr)) %>% ungroup() %>% mutate("se"=sd/sqrt(N))

expr_avg_df <- expr_cond_df %>% dplyr::select(id,condition,avg) %>% pivot_wider(names_from = condition, values_from = avg)
colnames(expr_avg_df)[-1]<-paste0(colnames(expr_avg_df)[-1],"_avg")

expr_se_df <- expr_cond_df %>% dplyr::select(id,condition,se) %>% pivot_wider(names_from = condition, values_from = se)
colnames(expr_se_df)[-1]<-paste0(colnames(expr_se_df)[-1],"_se")

expr_avgse_df<- suppressMessages(left_join(expr_avg_df,expr_se_df))
rownames(expr_avgse_df) <- expr_avgse_df$id
toPrint<-expr_avgse_df
colnames(toPrint)[1]<-"symbol"

n_pep_prot <- psm_anno_df %>% dplyr::group_by(symbol, .drop=FALSE) %>% dplyr::summarise("Num. Peptides"=n_distinct(row)) %>% ungroup()
toPrint_datGene<-dat_gene
toPrint_datGene$symbol<-rownames(dat_gene)

#Export table with SE e MEAN
df_to_save<-list("Normalized proteins matrix intensity"=left_join(n_pep_prot,toPrint_datGene),
                 "Normalized proteins intensity group by condition"=left_join(n_pep_prot,toPrint))
# writexl::write_xlsx(df_to_save,paste0(dirOutput,"normalised_intensity_table.xlsx"))
# remove(expr_mat,expr_l_df,expr_cond_df,expr_se_df,expr_avg_df,toPrint)


#Save peptides table
expr_mat<-dat_pep

expr_l_df <- expr_mat %>% mutate(id=rownames(expr_mat)) %>% pivot_longer(-id, names_to = "sample", values_to = "expr")
expr_l_df <- suppressMessages(left_join(expr_l_df,c_anno))

#Add Standard Deviation column and average
expr_cond_df<-expr_l_df %>% dplyr::group_by(condition,id) %>%
            dplyr::summarise(N=n(),avg=mean(expr),sd=sd(expr)) %>% ungroup() %>% mutate("se"=sd/sqrt(N))

expr_avg_pet_df <- expr_cond_df %>% dplyr::select(id,condition,avg) %>% pivot_wider(names_from = condition, values_from = avg)
colnames(expr_avg_pet_df)[-1]<-paste0(colnames(expr_avg_pet_df)[-1],"_avg")

expr_se_pet_df <- expr_cond_df %>% dplyr::select(id,condition,se) %>% pivot_wider(names_from = condition, values_from = se)
colnames(expr_se_pet_df)[-1]<-paste0(colnames(expr_se_pet_df)[-1],"_se")

expr_avgse_pep_df<- suppressMessages(left_join(expr_avg_pet_df,expr_se_pet_df))
rownames(expr_avgse_pep_df) <- expr_avgse_pep_df$id
toPrint<-expr_avgse_pep_df

#Export table with SE e MEAN
df_to_save<-list("Normalized proteins matrix intensity"=df_to_save$`Normalized proteins matrix intensity`,
                 "Normalized peptides matrix intensity"=merge(psm_peptide_table, dat_pep, by=0)[,-c(1)],
                 "Normalized proteins intensity group by condition"=df_to_save$`Normalized proteins intensity group by condition`,
                 "Normalized peptides intensity group by condition"=left_join(rownames_to_column(psm_peptide_table), 
                                                                              toPrint, by=c("rowname" = "id"))[,-c(1)])

writexl::write_xlsx(df_to_save,paste0(dirOutput,"normalised_intensity_table.xlsx"))
remove(expr_mat,expr_l_df,expr_cond_df,expr_se_pet_df,expr_avg_pet_df,toPrint)
```

#### 2.4. Normalized distributions
Box plot of the normalized and imputed data for each sample.

```{r norm_distributions, message=F, warning=F, echo=F}

pg_long_df <- dat_gene %>% mutate(id=as.factor(rownames(dat_gene))) %>% pivot_longer(-id, names_to = "sample", values_to = "intensity")
pg_long_df <- suppressMessages(left_join(pg_long_df,c_anno))


hs<-ggplot(data=pg_long_df, aes(x=sample,y=intensity,fill=sample,colour=sample))+
  # ggtitle("Normalized samples")+
  #geom_hline(yintercept = 0, lty=2)+
  coord_flip()+
  geom_violin(alpha=0.5,scale="width",trim = F,colour=NA)+
  geom_boxplot(alpha=1,fill="white",width=0.2,outlier.shape = NA, notch = F)+
  theme_bw()+
  theme(legend.position = "none",axis.title.y = element_blank())+
  labs(x = "Log2 Intensity") +
  scale_fill_manual(values=(colour_vec)) +
  scale_colour_manual(values=(colour_vec)) +
  theme(panel.grid.minor = element_blank(),panel.grid.major.y = element_blank())
  
hs

remove(hs, pg_long_df)
```


```{r mds_counts, warning=F, message=F, echo=F, results='asis'}
cat(" \n \n")
cat("#### 2.5. Multi-dimensional reduction analysis of samples {.tabset}  \n")
cat("Following are visualized the Multi-dimensional scaling analysis (MDSs) of the normalized data. The MDS in its simpler version is called PCA.Each sample has its own color described in the Input file, if it was not defined it is used the default palette coloring by condition. \n")
cat("Two MDSs are created, one based on the normalized protein intensities, tab **Proteins**. The second on the normalized peptide intensities, tab **Peptides** \n")
cat(" \n \n")

sample_data<-(t(dat_gene))
sample_distances <- dist(sample_data, method = "euclidean") 

mds_cmdscale <- as.data.frame(cmdscale(as.matrix(sample_distances)))
colnames(mds_cmdscale) <- c("MDS_1","MDS_2")
mds_cmdscale$sample <- rownames(mds_cmdscale)

mds_cmdscale <- suppressMessages(left_join(mds_cmdscale,c_anno))

#Take correct order of color
cc<-mds_cmdscale$color
names(cc)<-mds_cmdscale$condition

cmd<-ggplot(mds_cmdscale, aes(MDS_1, MDS_2, colour=condition)) + #plot according to the new coordinates
  geom_point(size=1.5,alpha=.9) + #colour=factor(df) to choose colour according to the condition
  geom_text_repel(aes(label=sample), size=0.15*bs, fontface = "bold",show.legend = F) + #geom_text_repel to add text within plot
  scale_colour_manual(values=cc[unique(names(cc))]) +
  theme_bw(base_size = bs, base_family = bf) + 
  theme(legend.position = "right", panel.grid.minor=element_blank())

cat("##### Proteins \n")
print(cmd)
cat(" \n \n")

suppressMessages(ggsave(paste0(dirOutput,"general_mds_proteins.pdf"), cmd, device=cairo_pdf, width = 4, height = 3, units = c("in")))

remove(sample_data,sample_distances,mds_cmdscale,cmd)

#Peptides
sample_data<-(t(dat_pep))
sample_distances <- dist(sample_data, method = "euclidean") 

mds_cmdscale <- as.data.frame(cmdscale(as.matrix(sample_distances)))
colnames(mds_cmdscale) <- c("MDS_1","MDS_2")
mds_cmdscale$sample <- rownames(mds_cmdscale)

mds_cmdscale <- suppressMessages(left_join(mds_cmdscale,c_anno))

#Take correct order of color
cc<-mds_cmdscale$color
names(cc)<-mds_cmdscale$condition

cmd<-ggplot(mds_cmdscale, aes(MDS_1, MDS_2, colour=condition)) + #plot according to the new coordinates
  geom_point(size=1.5,alpha=.9) + #colour=factor(df) to choose colour according to the condition
  geom_text_repel(aes(label=sample), size=0.15*bs, fontface = "bold",show.legend = F) + #geom_text_repel to add text within plot
  scale_colour_manual(values=cc[unique(names(cc))]) +
  theme_bw(base_size = bs, base_family = bf) + 
  theme(legend.position = "right", panel.grid.minor=element_blank())

cat("##### Peptides \n")
print(cmd)
cat(" \n \n")

suppressMessages(ggsave(paste0(dirOutput,"general_mds_peptides.pdf"), cmd, device=cairo_pdf, width = 4, height = 3, units = c("in")))

remove(sample_data,sample_distances,mds_cmdscale,cmd)
```

```{r batch_correction, warning=F, message=F, echo=F, eval=batch_corr_exe, results='asis'}
cat("#### 2.6. Batch correction {.tabset}  \n")
cat("A batch effect correction is applied to the datased using the proBatch package. \n")
cat("Two MDSs are created, one based on the corrected protein intensities, tab **Proteins**. The second on the corrected peptide intensities, tab **Peptides** \n")
cat(" \n \n")

batch_annotation <- c_anno
batch_annotation$order<-c(1:nrow(batch_annotation))

dat_pep_long <- matrix_to_long(dat_pep) %>% setnames("FullRunName","sample")
# loess_fit_df_pep <- adjust_batch_trend_df(dat_pep_long, batch_annotation, batch_col="ms_batch", sample_id_col="sample")
comBat_df_pep <- correct_with_ComBat_df(dat_pep_long, batch_annotation, batch_col="ms_batch", sample_id_col="sample")
dat_pep = long_to_matrix(comBat_df_pep, sample_id_col = "sample")

dat_gene_long <- matrix_to_long(dat_gene) %>% setnames("FullRunName","sample")
# loess_fit_df_gene <- adjust_batch_trend_df(dat_gene_long, batch_annotation, batch_col="ms_batch", sample_id_col="sample")
comBat_df_gene <- correct_with_ComBat_df(dat_gene_long, batch_annotation, batch_col="ms_batch", sample_id_col="sample")
dat_gene = long_to_matrix(comBat_df_gene, sample_id_col = "sample")

# setnames(c_anno,"FullRunName","sample")

###################
sample_data<-(t(dat_gene))
sample_distances <- dist(sample_data, method = "euclidean") 

mds_cmdscale <- as.data.frame(cmdscale(as.matrix(sample_distances)))
colnames(mds_cmdscale) <- c("MDS_1","MDS_2")
mds_cmdscale$sample <- rownames(mds_cmdscale)

mds_cmdscale <- suppressMessages(left_join(mds_cmdscale,c_anno))

#Take correct order of color
cc<-mds_cmdscale$color
names(cc)<-mds_cmdscale$condition

cmd<-ggplot(mds_cmdscale, aes(MDS_1, MDS_2, colour=condition)) + #plot according to the new coordinates
  geom_point(size=1.5,alpha=.9) + #colour=factor(df) to choose colour according to the condition
  geom_text_repel(aes(label=sample), size=0.15*bs, fontface = "bold",show.legend = F) + #geom_text_repel to add text within plot
  scale_colour_manual(values=cc[unique(names(cc))]) +
  theme_bw(base_size = bs, base_family = bf) + 
  theme(legend.position = "right", panel.grid.minor=element_blank())

cat("##### Proteins \n")
print(cmd)
cat(" \n \n")

suppressMessages(ggsave(paste0(dirOutput,"corrected_mds_proteins.pdf"), cmd, device=cairo_pdf, width = 4, height = 3, units = c("in")))

remove(sample_data,sample_distances,mds_cmdscale,cmd)

#Peptides
sample_data<-(t(dat_pep))
sample_distances <- dist(sample_data, method = "euclidean") 

mds_cmdscale <- as.data.frame(cmdscale(as.matrix(sample_distances)))
colnames(mds_cmdscale) <- c("MDS_1","MDS_2")
mds_cmdscale$sample <- rownames(mds_cmdscale)

mds_cmdscale <- suppressMessages(left_join(mds_cmdscale,c_anno))

#Take correct order of color
cc<-mds_cmdscale$color
names(cc)<-mds_cmdscale$condition

cmd<-ggplot(mds_cmdscale, aes(MDS_1, MDS_2, colour=condition)) + #plot according to the new coordinates
  geom_point(size=1.5,alpha=.9) + #colour=factor(df) to choose colour according to the condition
  geom_text_repel(aes(label=sample), size=0.15*bs, fontface = "bold",show.legend = F) + #geom_text_repel to add text within plot
  scale_colour_manual(values=cc[unique(names(cc))]) +
  theme_bw(base_size = bs, base_family = bf) + 
  theme(legend.position = "right", panel.grid.minor=element_blank())

cat("##### Peptides \n")
print(cmd)
cat(" \n \n")

suppressMessages(ggsave(paste0(dirOutput,"corrected_mds_peptides.pdf"), cmd, device=cairo_pdf, width = 4, height = 3, units = c("in")))

remove(sample_data,sample_distances,mds_cmdscale,cmd)

```

```{r boxplot_proteins, warning=F, message=F, echo=F, eval=boxplot_prot_exe, results='asis'}
prot_find<-unique(c(intersect(prot_boxplot,rownames(dat_gene)), 
                    intersect(str_to_title(prot_boxplot),rownames(dat_gene)),
                    intersect(str_to_upper(prot_boxplot),rownames(dat_gene)),
                    intersect(str_to_lower(prot_boxplot),rownames(dat_gene))))

if(length(prot_find)>0){
  cat("#### 2.7. Control boxplot proteins intensities {.tabset}  \n")
  cat("Boxplot figures for the control of the intensities of selected proteins. The list of proteins need to be written in the options file. For each protein shows the intensity group by condition. \n")
  cat("Selected proteins identified in the normalized list: \n")
  cat(paste(prot_find, sep = ", "))
  cat(" \n")
  cat(" \n \n")

  prot_intensity_long <- proBatch::matrix_to_long(dat_gene[prot_find,]) %>% 
                              data.table::setnames(new = c("Protein","Condition","Intensity"))
  prot_intensity_long$Condition <- c_anno[prot_intensity_long$Condition, "condition"]
  
  avg_df = expr_avg_df %>% dplyr::filter(id %in% prot_find) %>% dplyr::select(colnames(expr_avg_df)[-c(1)])
  se_df = expr_se_df %>% dplyr::filter(id %in% prot_find) %>% dplyr::select(colnames(expr_se_df)[-c(1)])
  prot_avg_long <- proBatch::matrix_to_long(avg_df) %>% data.table::setnames(new = c("Protein","Condition","avg"))
  prot_avg_long$Protein<-(expr_avg_df %>% dplyr::filter(id %in% prot_find))$id[as.integer(prot_avg_long$Protein)]
  prot_se_long <- proBatch::matrix_to_long(se_df) %>% data.table::setnames(new = c("Protein","Condition","se"))
  prot_se_long$Protein<-(expr_se_df %>% dplyr::filter(id %in% prot_find))$id[as.integer(prot_se_long$Protein)]
  prot_avg_se_long <- cbind(prot_avg_long,prot_se_long[3])
  prot_avg_se_long$Condition <- str_replace(prot_avg_se_long$Condition, "_avg", "")
  
  g<-ggplot(prot_avg_se_long,aes(Condition,avg,fill=Condition,colour=Condition))+
    geom_crossbar(aes(ymin=avg,ymax=avg),position = "dodge",width=.8,alpha=.9,fatten=1.5)+
    geom_errorbar(aes(ymin=(avg-se), ymax=(avg+se)), width=.4,position=position_dodge(),show.legend=F,alpha=.8)+
    geom_quasirandom(data=prot_intensity_long, aes(Condition,Intensity), alpha=.7,width=.1,shape=16,size=0.11*bs)+
    scale_fill_manual(name="Condition",values=cc) +
    scale_colour_manual(name="Condition",values=cc)+
    # scale_y_continuous(limits=c(0,NA),expand = expand_scale(mult = c(.1, .25)))+
  
    theme_bw(base_size = bs, base_family = bf) +
    theme(axis.title.x=element_blank()) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1, colour=cc))+
    theme(panel.grid.major.x=element_blank(),
          panel.grid.minor.y=element_blank())+
    theme(legend.text = element_text(size = 0.7*bs),
          legend.key.size = unit((0.015*bs),"in"),
          legend.position="none",
          legend.title=element_blank(),
          legend.background = element_rect(fill = NA))+
   
    theme(strip.text=element_text(colour="white",face="bold"))+
    theme(panel.border=element_rect(colour=c("grey40"),size=0.03*bs))+
    theme(strip.background=element_rect(fill="grey40",colour="grey40",size=0.03*bs))+
    theme(plot.title = element_text(hjust = 0.5))+
    facet_wrap(~Protein, scales = "free",ncol = 4)+
    labs(y="Abundance")
  print(g)


  suppressMessages(ggsave(paste0(dirOutput,"boxplot_control_intensity.pdf"), g, device=cairo_pdf, width = 5, height = 5, units = c("in")))
  
}else{
  cat("##### Selected proteins in options file do not find in the normalised matrix. Check the spell of the proteins. \n")
  cat(" \n \n")
}
```

## 3. Identification of Differentially Expressed Proteins and Peptides

Differential expression analysis for proteins is performed with the DEqMS package. DEqMS is developed on top of Limma, but the method estimates different prior variances for proteins quantified by different number of PSMs/peptides, therefore achieving better accuracy. 
This methodology can not be applied to the peptides, so, the classical Limma method is used.
In the plot are represented graphically the number of DEPs for each contrast. Tab **Proteins** for the differentially expressed proteins, **Peptides** for the number of the differentially expressed peptides.


```{r diff_analysis,  warning=F, message=F, echo=F}
message("Differentiation analysis in progress...")
#Execution of differentiation analysis using DEqMS for proteins
psm_count_table = as.data.frame(table(dat_psm$symbol))
deps_df <- limmafnc("PROT",c_anno,dat_gene,psm_count_table,formule_contrast,expr_avgse_df,signal_thr,fc_thr, pval_thr)
deps_l_df<-deps_df$degs_l_df
deps_w_df<-deps_df$degs_w_df
toPrint<-deps_w_df
colnames(toPrint)[1]<-"GeneSymbol"

#Execution of differentiation analysis using DEqMS for proteins
psm_count_table<-as.data.frame(table(rownames(dat_pep)))
deps_pep_df <- limmafnc("PEP",c_anno,dat_pep,psm_count_table,formule_contrast,expr_avgse_pep_df,signal_thr,fc_thr, pval_thr)
deps_pep_l_df<-deps_pep_df$degs_l_df
deps_pep_w_df<-deps_pep_df$degs_w_df
toPrint_pep<-deps_pep_w_df
df_to_save<-list("DEPs_proteins"=toPrint,
                 "DEPs_peptides"=left_join(rownames_to_column(psm_peptide_table), toPrint_pep, by=c("rowname" = "id"))[,-c(1)])
writexl::write_xlsx(df_to_save,paste0(dirOutput,"DEPs_table.xlsx"))

remove(deps_df,toPrint,deps_pep_df,toPrint_pep,df_to_save)
message("Differentation analysis DONE.")
```

```{r number_deps, warning=FALSE, echo=FALSE, results='asis'}

# Plot DEPs barplot (as a double barplot)

plot_DEPs<-function(DEPs_df, # dataframe containing DEPs result
                    contrast_col="comp", # Name of the column with contrast names
                    class_col="class", # Name of the column with classes (+,-,= are expected)
                    color_vec=c("#EAB364", "#84A298") # Colors for Up and Down regulated
                     ){
  res_mat<-DEPs_df %>% group_by(get(contrast_col),get(class_col),.drop=FALSE) %>% summarise("Number"=n()) %>% ungroup() 
  # get() to return the value of variable 
  # in this way we can assign the name of the column to a variable and keep the same script changing every time only the name of the column
  # .drop=FALSE to keep the comparisons with zero-value classes (for example in the fdr df)
  colnames(res_mat)<-c("Contrast","Class","Number")
  res_df<-res_mat %>% filter(Class!="=")
  
  res_df$Class<-factor(res_df$Class,levels=c("+","-"))
  res_df$Contrast<-as.factor(res_df$Contrast)
  res_df$Number<-as.numeric(res_df$Number)
  res_df$Number[which(res_df$Class=="-")]<-res_df$Number[which(res_df$Class=="-")]*(-1) 
  # *(-1) to assign to the total number of down-reg a - sign and put them on the left of the plot
  
  #Change ratio based on the number of comp
  n_comp<-length(unique(res_df$Contrast))
  ratio<-1/((5/n_comp)+0.25)
  bs<-11
  bap<-ggplot(res_df,aes(x=Contrast,y=Number,fill=Class)) +
    scale_fill_manual(name="",values=color_vec,labels=c("Up-Regulated","Down-Regulated"))+
    geom_bar(stat="identity",position=position_nudge(),width=0.6,alpha=0.9) +
    geom_text(aes(x=Contrast, y=Number, label=abs(Number), hjust=ifelse(res_df$Number>=0,1.1,-0.1)), colour="white",size=bs*0.4,fontface="plain") +
    theme_bw(base_size = bs) +
    theme(axis.title.y=element_blank()) +
    theme(panel.border=element_blank()) +
    theme(axis.ticks=element_blank()) +
    theme(panel.grid=element_blank()) +
    theme(aspect.ratio = ratio) +
    geom_hline(yintercept=0,col="grey30",lty=1,lwd=0.05*bs)+
    theme(axis.text.y=element_text(colour="grey20")) +
    theme(axis.text.x=element_blank()) +
    theme(legend.position="bottom") +
    #theme(legend.justification=c(1.05,1.05), legend.position=c(1,0.07)) +
    theme(legend.title=element_blank())+
    coord_flip()+
    theme(legend.text = element_text(size = bs),
          legend.key.size = unit((0.012*bs),"in"))+
    labs(y="Number of DEPs")
  return(bap)
}

cat(" \n \n")
cat("#### Number of DEPs fro each comparison {.tabset}  \n")
#Execute only if found at least one gene differentiated
if(deps_l_df %>% dplyr::filter(class!="=") %>% nrow() > 0){
  pDEPs<-plot_DEPs(deps_l_df,
                    contrast_col="comp", 
                    class_col="class",
                    color_vec=c("grey40", "grey60"))
  cat("##### Proteins \n")
  print(pDEPs)
  cat(" \n \n")
  
  suppressMessages(ggsave(paste0(dirOutput,"DEPs_proteins_count_barplot.pdf"), pDEPs, device=cairo_pdf, units = c("in")))
  remove(pDEPs,cmd,mds_cmdscale)
}else{
  message("No gene differentiated discovered.")
  cat("No gene differentiated discovered.")
}
#Peptides
if(deps_pep_l_df %>% dplyr::filter(class!="=") %>% nrow() > 0){
  cat(" \n \n")
  pDEPs<-plot_DEPs(deps_pep_l_df,
                    contrast_col="comp", 
                    class_col="class",
                    color_vec=c("grey40", "grey60"))
  cat("##### Peptides \n")
  print(pDEPs)
  cat(" \n \n")
  
  suppressMessages(ggsave(paste0(dirOutput,"DEPs_peptides_count_barplot.pdf"), pDEPs, device=cairo_pdf, units = c("in")))
  remove(pDEPs,cmd,mds_cmdscale)
}else{
  message("No gene differentiated discovered.")
  cat("No gene differentiated discovered.")
}

```

```{r volcano_protein, fig.height=6, fig.width=6, warning=FALSE, echo=FALSE, results='asis'}
cat(" \n \n")
if(deps_l_df %>% dplyr::filter(class!="=") %>% nrow() > 0){
  
  cat("#### 3.1.1. Volcano plot of Proteins {.tabset}  \n")
  cat("In the tabs below, the figures represent the volcano plots of differentially expressed proteins. Each condition has its own tab with the correspoding figure. \n")
  cat("The line in the volcano are the threshold set in the **options** file. \n")
  cat("Class identify the state of the protein. \"+\" are the Up-regulated DEPs, \"-\" are the Down-regulated DEPs, \"=\" are the non-significative DEPs. The tool writes the labels of the proteins which can fit in the graph. \n")
  cat(" \n \n")
  
  for(comp in unique(deps_l_df$comp)){
    input_df <- deps_l_df[which(deps_l_df$comp == comp),]
    labels_up <- (input_df %>% filter(class == '+'))$id
    labels_down <- (input_df %>% filter(class == '-'))$id
    
    cmd<-ggplot(data=input_df, aes(x=log2_FC, y=-log10(p_val), col=class, label=id)) +
      geom_point(pch=20,cex=2) +
      geom_hline(yintercept=-log10(pval_thr), col="black") +
      geom_vline(xintercept=c(-fc_thr, fc_thr), col="black") +
      # geom_text_repel(data = input_df %>% filter(id %in% labels),color="white") +
      geom_label_repel(data = input_df %>% filter(id %in% labels_up), label.padding = unit(0.1, "lines"), color="black",xlim = c(1.5, NA)) +
      geom_label_repel(data = input_df %>% filter(id %in% labels_down), label.padding = unit(0.1, "lines"), color="black",xlim = c(NA, -1.5)) +
      ggtitle(paste0("Volcano Plot of ",comp)) +
      scale_color_manual(values=c(color_constrast[comp,"color_constrast"], color_constrast[comp,"color_constrast_dark"], "grey70")) +
      scale_x_continuous(limits = c(-max(abs(input_df$log2_FC)), max(abs(input_df$log2_FC)))) +
      theme_bw(base_size = bs, base_family = bf)
    
    cat("##### ",comp," \n")
    print(cmd)
    cat(" \n \n")
  }
}
```

```{r volcano_pep, fig.height=6, fig.width=6, warning=FALSE, echo=FALSE, results='asis'}
cat(" \n \n")
if(deps_pep_l_df %>% dplyr::filter(class!="=") %>% nrow() > 0){
  
  cat("#### 3.1.2. Volcano plot of Peptides {.tabset}  \n")
  cat("In the tabs below, the figures represent the volcano plots of differentially expressed peptides. Each condition has its own tab with the correspoding figure. \n The line in the volcano are the threshold set in the **options** file. \n Class identify the state of the peptides. \"+\" are the Up-regulated DEPs, \"-\" are the Down-regulated DEPs, \"=\" are the non-significative DEPs. The tool writes the labels of the peptides which can fit in the graph. \n")
  cat(" \n \n")
  
  for(comp in unique(deps_pep_l_df$comp)){
    input_df <- na.omit((deps_pep_l_df[which(deps_pep_l_df$comp == comp),])[order(deps_pep_l_df$id),])
    pos<-lapply(str_match_all(str_split(psm_peptide_table$`Position in Master Proteins`,";",n=1),"\\[(.*?)\\]"), function(x){x[,1][1]})
    labels<-str_c(psm_peptide_table$GeneName,pos,sep="_")
    input_df$id<-labels
    labels_up <- (input_df %>% filter(class == '+'))$id
    labels_down <- (input_df %>% filter(class == '-'))$id
    
    cmd<-ggplot(data=input_df, aes(x=log2_FC, y=-log10(p_val), col=class, label=id)) +
      geom_point(pch=20,cex=2) +
      geom_hline(yintercept=-log10(pval_thr), col="black") +
      geom_vline(xintercept=c(-fc_thr, fc_thr), col="black") +
      # geom_text_repel(data = input_df %>% filter(id %in% labels),color="white") +
      geom_label_repel(data = input_df %>% filter(id %in% labels_up, class == '+'), label.padding = unit(0.1, "lines"), color="black",xlim = c(1.5, NA)) +
      geom_label_repel(data = input_df %>% filter(id %in% labels_down, class == '-'), label.padding = unit(0.1, "lines"), color="black",xlim = c(NA, -1.5)) +
      ggtitle(paste0("Volcano Plot of ",comp)) +
      scale_color_manual(values=c(color_constrast[comp,"color_constrast"], color_constrast[comp,"color_constrast_dark"], "grey70")) +
      scale_x_continuous(limits = c(-max(abs(input_df$log2_FC)), max(abs(input_df$log2_FC)))) +
      theme_bw(base_size = bs, base_family = bf)
    
    cat("##### ",comp," \n")
    print(cmd)
    cat(" \n \n")
  }
}
```

```{r PCA_deps, warning=F, message=F, echo=F, eval=T, results='asis'}
if(deps_l_df %>% dplyr::filter(class!="=") %>% nrow() > 0){
  cat("#### 3.2.1. Multi-dimensional reduction analysis of samples based on up-/down-regulated proteins \n")
  cat("The below figure is the MDS analysis on the significative differentially expressed proteins. \n \n")
  #Select only genes that are differentiated and print PCA of samples
  deps_vec_all<- deps_l_df %>% dplyr::filter(class!="=") %>% 
  pull(id) %>% unique() %>% sort()
  
  # sample_data<-t(dat_gene[toupper(deps_vec_all),]) 
  sample_data<-t(dat_gene[(deps_vec_all),]) 
      
  sample_distances <- dist(sample_data, method = "euclidean") 
    
  mds_cmdscale <- as.data.frame(cmdscale(as.matrix(sample_distances)))
  colnames(mds_cmdscale) <- c("MDS_1","MDS_2")
  mds_cmdscale$sample <- rownames(mds_cmdscale)
    
  mds_cmdscale <- suppressMessages(left_join(mds_cmdscale,c_anno))
  #Take correct order of color
  cc<-mds_cmdscale$color
  names(cc)<-mds_cmdscale$condition
  cmd<-ggplot(mds_cmdscale, aes(MDS_1, MDS_2, colour=condition)) + #plot according to the new coordinates
    # ggtitle("samples")+
    geom_point(size=1.5,alpha=.9) + #colour=factor(df) to choose colour according to the condition
    geom_text_repel(aes(label=sample), size=0.15*bs, fontface = "bold",show.legend = F) + #geom_text_repel to add text within plot
    scale_colour_manual(values=cc[unique(names(cc))]) +
    theme_bw(base_size = bs, base_family = bf) + 
    theme(legend.position = "right", panel.grid.minor=element_blank())
  
  print(cmd)
  
  ggsave(paste0(dirOutput,"DEPs_proteins_mds.pdf"), cmd, device=cairo_pdf, width = 4, height = 3, units = c("in"))
  
}
```

```{r PCA_deps_pep, warning=F, message=F, echo=F, eval=T, results='asis'}
if(deps_pep_l_df %>% dplyr::filter(class!="=") %>% nrow() > 0){
  cat("#### 3.2.2. Multi-dimensional reduction analysis of samples based on up-/down-regulated peptides \n")
  cat("The below figure is the MDS analysis on the significative differentially expressed peptides. \n \n")
  #Select only genes that are differentiated and print PCA of samples
  deps_vec_all<- deps_pep_l_df %>% dplyr::filter(class!="=") %>% 
  pull(id) %>% unique() %>% sort()
  
  # sample_data<-t(dat_gene[toupper(deps_vec_all),]) 
  sample_data<-t(dat_pep[(deps_vec_all),]) 
      
  sample_distances <- dist(sample_data, method = "euclidean") 
    
  mds_cmdscale <- as.data.frame(cmdscale(as.matrix(sample_distances)))
  colnames(mds_cmdscale) <- c("MDS_1","MDS_2")
  mds_cmdscale$sample <- rownames(mds_cmdscale)
    
  mds_cmdscale <- suppressMessages(left_join(mds_cmdscale,c_anno))
  #Take correct order of color
  cc<-mds_cmdscale$color
  names(cc)<-mds_cmdscale$condition
  cmd<-ggplot(mds_cmdscale, aes(MDS_1, MDS_2, colour=condition)) + #plot according to the new coordinates
    # ggtitle("samples")+
    geom_point(size=1.5,alpha=.9) + #colour=factor(df) to choose colour according to the condition
    geom_text_repel(aes(label=sample), size=0.15*bs, fontface = "bold",show.legend = F) + #geom_text_repel to add text within plot
    scale_colour_manual(values=cc[unique(names(cc))]) +
    theme_bw(base_size = bs, base_family = bf) + 
    theme(legend.position = "right", panel.grid.minor=element_blank())
  
  print(cmd)
  
  ggsave(paste0(dirOutput,"DEPs_peptides_mds.pdf"), cmd, device=cairo_pdf, width = 4, height = 3, units = c("in"))
  
}
```



```{r enrichment,message=F, warning=F,echo=F, eval=run_enrich, results='asis'}
cat(" \n \n")
cat("## 4. Functional analysis of Differentially Expressed Proteins \n") 
cat("The enrichment analysis is done by [EnrichR](https://maayanlab.cloud/Enrichr/). It is a web-base application which integrates a large number of data sets that permit to obtain information about many categories. EnrichR organises its hundreds of data sets in 8 sections: Transcription, Pathways, Ontologies, Diseases/Drugs, Cell Types, Misc, Legacy, and Crowd. All the libraries of EnrichR are queried to provide the most complete research. \n")
cat(" Each comparison is analysed through 3 sets of proteins: the Up-regulated (called **UP**), the Down-regulated (called **DOWN**), and the set (called **ALL**) obtained by the merge of Up- and Down-regulated proteins. \n")
def_val_height=  knitr::opts_chunk$get("fig.height")
message("Enrichment step with EnrichR in proress...")
doNextChunk <- tryCatch({
  #Perform enriched analysis with EnrichR
  enr_df<-enrichRfnc(deps_l_df)
  #Save in RData for possible further analysis
  save(enr_df, file=paste0(dirOutput,"DEPs_enrichment_data.RData"))

  writexl::write_xlsx((enr_df %>% dplyr::filter(p_value<=pval_enrich_thr,overlap_size>=overlap_size_enrich_thr)),
                      path = paste0(dirOutput,"DEPs_enrichment_table.xlsx"))
  dir.create(file.path(dirOutput,"enrich/"), showWarnings = FALSE)
  dirOutput_enr=paste0(dirOutput,"enrich/")
  TRUE
},
error=function(cond){
  print("WARNING: Not possible to continue with the enrichment. No protein up- or down-regulated.\n")
  FALSE
})
message("Enrichment DONE.")
```

```{r enrichment_plots, echo=F, message=FALSE, warning=FALSE, eval = doNextChunk, results='asis'}

if(length(enrich_filter_DBs) > 0){
  cat("#### 4.1. Enrichment of DEPs lists filtered by databases in options file.\n")
  cat("The figure is done by the UP (up-regulated proteins) and the DOWN (down-regulated proteins) analysis. The graph visualizes only the significant enriched terms of the pre-selected data sets. \n")
  cat("The dimension of the point is given by the Odds Ratio in log2 scale. The filling is based on the significativity of the P.Value against the selected threshold. \n")
  
  #Filter the DEPs taking only the ones from the desired DBs
  annos<- enrich_filter_DBs
  
  enr_sele_names <- enr_df %>% 
    dplyr::filter(anno_class %in% annos,
                  overlap_size>=overlap_size_enrich_thr,
                  anno_size<=10000,
                  p_value<pval_enrich_thr) %>% 
    dplyr::arrange(-overlap_size) %>% dplyr::group_by(input_name,anno_class) %>% slice_head(n=5) %>% 
    ungroup() %>% dplyr::select(anno_class,anno_name)
  
  enr_sele_df<-suppressMessages(inner_join(enr_df,enr_sele_names))
  
  #Select the results divided in up and down
  enr_sele_df <- enr_sele_df %>% dplyr::filter(!str_ends(input_name,"_all"))
  knitr::opts_chunk$set(fig.height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3))
}
```

```{r enrichment_plots_2, echo=F, message=FALSE, warning=FALSE, eval = doNextChunk, results='hide'}

if(length(enrich_filter_DBs) > 0){
  #Print the plot
  # col_vec<-color_base[sort(rep(c(1:length(color_base)),2))]
  # if(length(as.vector(t(color_constrast)))>0){
  col_vec<-as.vector(t(color_constrast))
  names(col_vec)<-unique(enr_sele_df$input_name)
  # }
  el<- enrichment_dotmatrix(enr_sele_df,
                           size_col="log2_OR",
                           color_col="input_name",
                           color_vec = col_vec,
                           # color_vec=color_base[sort(rep(c(1:length(color_base)),2))],
                           char_max=60)
  
  tryCatch({
    print(el +guides(color=F))
    message(paste0(dirOutput_enr,"enrichment_database_filter.pdf"))
    ggsave(paste0(dirOutput_enr,"enrichment_database_filter.pdf"), el, device=cairo_pdf, width = 7, height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3), units = c("in"))
  },
  error=function(cond){
    print("Error: No enriched term found with current parameters\n")
  })
}

```

```{r enrichment_plots_all, echo=F, message=F, warning=F, eval = doNextChunk, results='asis'}

if(length(enrich_filter_DBs) > 0){
  cat("#### 4.2. Enrichment of DEPs lists filtered by databases in options file.\n")
  cat("The figure is done by the ALL (join of up- and down-regulated proteins) analysis. The graph visualizes only the significant enriched terms of the pre-selected data sets. \n")
  cat(" The dimension of the point is given by the Odds Ratio in log2 scale. The filling is based on the significativity of the P.Value against the selected threshold. \n")
  #Select DEPs in the DBs of interest
  annos<- enrich_filter_DBs
  
  enr_sele_names <- enr_df %>% 
    dplyr::filter(anno_class %in% annos,
                  overlap_size>=overlap_size_enrich_thr,
                  anno_size<=10000,
                  p_value<pval_enrich_thr) %>% 
    dplyr::arrange(-overlap_size) %>% dplyr::group_by(input_name,anno_class) %>% slice_head(n=5) %>% 
    ungroup() %>% dplyr::select(anno_class,anno_name)
  
  enr_sele_df<-suppressMessages(inner_join(enr_df,enr_sele_names))
  
  #Select the ALL column (UP union DOWN)
  enr_sele_df <- enr_sele_df %>% dplyr::filter(str_ends(input_name,"_all"))
  knitr::opts_chunk$set(fig.height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3))
}
```

```{r enrichment_plots_all_2, echo=F, message=FALSE, warning=FALSE, eval = doNextChunk, results='hide'}

if(length(enrich_filter_DBs) > 0){
    # col_vec<-color_base
    # if(length(as.vector(t(color_constrast)))>0){
    col_vec<-as.vector((color_constrast))
    names(col_vec)<-unique(enr_sele_df$input_name)
    # }
    el<- enrichment_dotmatrix(enr_sele_df,
                           size_col="log2_OR",
                           color_col="input_name",
                           color_vec=col_vec,
                           char_max=60)
  
  tryCatch({
    print(el +guides(color=F))
    ggsave(paste0(dirOutput_enr,"enrichment_database_filter_all.pdf"), el, device=cairo_pdf, width = 7, height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3), units = c("in"))
  },
  error=function(cond){
    print("Error: No enriched term found with current parameters\n")
  })
}
```

```{r enrichment_plots_terms, echo=F, message=F, warning=F, eval = doNextChunk, results='asis'}

if(length(enrich_filter_term) > 0){
  cat("#### 4.3. Enrichment of DEPs lists filtered by terms in options file. \n ")
  cat("The figure is done by the UP (up-regulated proteins) and the DOWN (down-regulated proteins) analysis. The graph visualizes only the significant enriched terms which contains the selected key-words. \n")
  cat(" The dimension of the point is given by the Odds Ratio in log2 scale. The filling is based on the significativity of the P.Value against the selected threshold. \n")
  cat(" \n")
  cat("**List of terms to search:**")
  cat(" \n")
  cat(paste(enrich_filter_term, collapse=", "))
  
  lookup_words<-enrich_filter_term
  
  enr_sele_names<-NULL
  
  #For each patter we want, search in each DEPs name to find it
  for(word in lookup_words){
    matches <- enr_df %>% dplyr::filter(p_value<pval_enrich_thr,
                                        overlap_size>=overlap_size_enrich_thr) %>% dplyr::select(anno_class,anno_name)
    matches <- matches[grepl(paste0("\\b",word,"\\b"),matches$anno_name,ignore.case=TRUE),] 
    enr_sele_names<-rbind(enr_sele_names,matches) %>% unique()
  }
  
  enr_sele_df<-suppressMessages(inner_join(enr_df,enr_sele_names))
  # write_tsv(enr_sele_df ,file=paste(dirOutput,dirOutput_ev,"enrich_myc_v1_all.txt",sep=""))
  
  enr_sele_df <- enr_sele_df %>% dplyr::filter(!str_ends(input_name,"_all"))
  knitr::opts_chunk$set(fig.height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3))
}
```

```{r enrichment_plots_terms_2, echo=F, message=FALSE, warning=FALSE, eval = doNextChunk, results='hide'}

if(length(enrich_filter_term) > 0){
  # col_vec<-color_base[sort(rep(c(1:length(color_base)),2))]
  # if(length(as.vector(t(color_constrast)))>0){
  col_vec<-as.vector(t(color_constrast))
  names(col_vec)<-unique(enr_sele_df$input_name)
  # }  
  el<- enrichment_dotmatrix(enr_sele_df,
                         size_col="log2_OR",
                         color_col="input_name",
                         color_vec = col_vec,
                         # color_vec=color_base[sort(rep(c(1:length(color_base)),2))],
                         char_max=60) + guides(color=F)
  tryCatch({
    print(el) 
  
    ggsave(paste0(dirOutput_enr,"enrichment_term_filter.pdf"), el, device=cairo_pdf, width = 7, height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3), units = c("in"))
  },
  error=function(cond){
    print("Error: No enriched term found with current parameters\n")
  })
  
}
```

```{r enrichment_plots_all_terms, echo=F, message=F, warning=F, eval = doNextChunk, results='asis'}

if(length(enrich_filter_term) > 0){
  cat("#### 4.4. Enrichment of DEPs lists filtered by terms in options file. \n")
  cat("The figure is done by the ALL (join of up- and down-regulated proteins) analysis. The graph visualizes only the significant enriched terms which contains the selected key-words. \n")
  cat(" The dimension of the point is given by the Odds Ratio in log2 scale. The filling is based on the significativity of the P.Value against the selected threshold. \n")
  cat(" \n")
  cat("**List of terms to search:**")
  cat(" \n")
  cat(paste(enrich_filter_term, collapse=", "))
  
  lookup_words<-enrich_filter_term
  
  enr_sele_names<-NULL
  #For each pattern search in DEPs name
  for(word in lookup_words){
    matches <- enr_df %>% dplyr::filter(p_value<pval_enrich_thr,
                                        overlap_size>=overlap_size_enrich_thr) %>% dplyr::select(anno_class,anno_name)
    matches <- matches[grepl(paste0("\\b",word,"\\b"),matches$anno_name,ignore.case=TRUE),] 
    enr_sele_names<-rbind(enr_sele_names,matches) %>% unique()
  }
  
  enr_sele_df<-suppressMessages(inner_join(enr_df,enr_sele_names))
  # write_tsv(enr_sele_df ,file=paste(dirOutput,dirOutput_ev,"enrich_myc_v1_all.txt",sep=""))

  enr_sele_df <- enr_sele_df %>% dplyr::filter(str_ends(input_name,"_all"))
  knitr::opts_chunk$set(fig.height = max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3))
}
```

```{r enrichment_plots_all_terms_2, echo=F, message=FALSE, warning=FALSE, eval = doNextChunk, results='hide'}

if(length(enrich_filter_term) > 0){
  # col_vec<-color_base
  #   if(length(as.vector(t(color_constrast)))>0){
  col_vec<-as.vector((color_constrast))
  names(col_vec)<-unique(enr_sele_df$input_name)
    # }
  el<- enrichment_dotmatrix(enr_sele_df,
                         size_col="log2_OR",
                         color_col="input_name",
                         color_vec=col_vec,
                         char_max=60) + guides(color=F)
  
  tryCatch({
    print(el) 
  
    ggsave(paste0(dirOutput_enr,"enrichment_term_filter_all.pdf"), el, device=cairo_pdf, width = 7, height =
             max(min(20,length(unique(enr_sele_df$anno_name))*0.5),3), units = c("in"))
  },
  error=function(cond){
    print("Error: No enriched term found with current parameters\n")
  })
}
```

 

```{r selection_genes, echo=F, message=F, warning=F, eval=run_STRING, results='asis'}
cat(" \n \n")
cat("## 5. Differetially Expressed Proteins in STRINGdb network \n")
cat("Protein-protein interactions (PPIs) is crucial for understanding cell physiology in different states. Also, knowledge from PPIs can be used (EMBL-EBI, 2022) for characterize unknown proteins, study pathways, characterize relationship between proteins. The pipeline uses [STRING](https://string-db.org/) to execute the PPI analysis. The database is prefilter removing the edges with low strength. The iteration searchs the interactions between DEPs in the database. Discovered all the connections, the clusters are identified. \n")
#Reset of height of plots
if(run_enrich){knitr::opts_chunk$set(fig.height = def_val_height)}

#Select up- down- regualted genes, if 0 genes are find, the all network analysis is skipped
g_sel_comp<-list()
doComp<-list()
for (comp in unique(deps_l_df$comp)){
  genes<-deps_l_df[(deps_l_df$class != "=" & deps_l_df$comp == comp), "id"]
  if(length(genes)>0){g_sel_comp[[comp]]<-genes}else{g_sel_comp[[comp]]<-list()}
  if(length(g_sel_comp[[comp]])!=0){doComp[comp]=TRUE}
}
if(length(doComp)==0 | all(doComp==F)){
  doNextChunk_STRING<-FALSE
  print("WARNING: Not possible to continue with the STRING network evaluation. No protein up- or down-regulated.\n")
}else{
  doNextChunk_STRING<-TRUE
  dir.create(file.path(dirOutput,"network/"), showWarnings = FALSE)
  dirOutput_net=paste0(dirOutput,"network/")
}
```

```{r network_string_communities, echo=F, message=F, warning=F, eval = doNextChunk_STRING, results='asis'}
message("Build PPI network with STRING in progress...")

#Read STRINGdb edges for human
#Human id --> 9606
#Mouse id --> 10090
string_db <- STRINGdb$new(version="11.5", species=taxonomy_NCBI, score_threshold=500, input_directory="")
gene_name<-unlist(g_sel_comp) %>% unique() %>% as.data.frame()
colnames(gene_name)<-c("gene_id")
string_mapped <- string_db$map(gene_name, "gene_id", removeUnmappedRows = TRUE)
links_string <- string_db$get_interactions(string_mapped$STRING_id)
links_string$from <- string_mapped$gene_id[match(links_string$from,string_mapped$STRING_id)]
links_string$to <- string_mapped$gene_id[match(links_string$to,string_mapped$STRING_id)]
string_gene_df<-data_frame("gene1"=links_string$from,"gene2"=links_string$to,"weigth"=links_string$combined_score)

if(nrow(string_gene_df) > 0){
  # cat(" \n")
  # cat("#### 5.1. Number of proteins in each communities discovered \n")
  # cat("Each image is a different contrast analysis \n")
  data_comms_df<-list()
  data_comms_list<-list()
  data_dt_links<-list()
  colour_vector<-list()
  #Cycle for each contrast
  for (comp in names(doComp)){
    # params ----
    thr_score <- 150 # define max strength of interaction
    
    #Select genes that are up or down regulated
    g_sel <- g_sel_comp[[comp]]
    if(unique(!grepl("[[:lower:]]",string_gene_df$gene1))){
      g_sel<-toupper(g_sel)
    }else{
      g_sel<-str_to_title(g_sel)
    }
    #Find communities in STRING db
    comm <- suppressMessages(find_communities(g_sel, thr_score,string_gene_df))
    i_comms_df <- comm[["i_comms_df"]]
    data_comms_df[[comp]]<-i_comms_df
    i_comms_list <- comm[["i_comms_list"]]
    data_comms_list[[comp]]<-i_comms_list
    dt_links <- comm[["dt_links"]]
    data_dt_links[[comp]]<-dt_links
    
    # colour_vector[[comp]] <- c(c(brewer.pal(n = 8, name = "Dark2")), "#680000", "#ae0001", "#eb8c00", "#680000", "#001080", "#999999", "#434343")
    colour_vector[[comp]] <- c(c(brewer.pal(n = 8, name = "Dark2")), color_base)
    colour_vector[[comp]] <- colour_vector[[comp]][1:length(i_comms_list)]
    
    # nc <- ggplot(i_comms_df, aes(x=comm_n, fill=comm_n)) +
    #   geom_bar(alpha=0.9, colour="white", width=0.4)+
    #   scale_x_discrete(drop=FALSE)+
    #   scale_fill_manual(values=colour_vector[[comp]],drop=FALSE)+
    #   theme_bw(base_size = bs) +
    #   labs(x = "Community", y = "# genes") +
    #   theme(legend.position = "none", panel.grid.major.x = element_blank(),
    #         panel.grid.minor.x = element_blank(), panel.grid.minor.y = element_blank())+
    #   ggtitle("Number of genes in each community", subtitle = paste0("Before Filtering of ",comp))
    
    # print(nc)
    cat("\n")
    # ggsave(paste0(dirOutput_net,comp,"_community_number_gene_prefilter.pdf"), nc, device=cairo_pdf, width = 4, height = 4, units = c("in"))
  
  } 
  
  i_comms_df<-data_comms_df
  i_comms_list<-data_comms_list
  dt_links<-data_dt_links
  
  message("Network DONE.")
}else{
  print("WARNING: No strong interaction detected beetween the proteins. Usually too few proteins.")
  doNextChunk_STRING <- FALSE
  message("Network incomplete.")
}
```

```{r network_string, echo=F, message=F, warning=F, eval = doNextChunk_STRING, fig.width=12, results='asis'}
message("Plotting STRING network in progress...")
cat("#### 5.1. Number of proteins in each communities discovered  {.tabset}  \n")
cat("The bar plots below describe the communities discovered in each network. THe heigth of the bar is the number of gene in that community. Each tab is a different contrast analysis. ")
cat(" \n \n")


#Multicore parallelizatio of plot
ncores <- 2
cluster_ext <- makeCluster(ncores, type = "SOCK")
registerDoParallel(cl = cluster_ext)

for (comp in names(doComp)){
  scr_thr <-  200# visual
  links <- 1
  # Filter genes and links ----
  # net_genes <- toupper(g_sel_comp[[comp]])
  net_genes <- g_sel_comp[[comp]]
  if(unique(!grepl("[[:lower:]]",string_gene_df$gene1))){
      net_genes<-toupper(net_genes)
    }else{
      net_genes<-str_to_title(net_genes)
    }
  net_edges <- subset(dt_links[[comp]], from%in%net_genes & to%in%net_genes) #select genes with weight>scr_thr
  net_edges<-net_edges[which(net_edges$weight > mean(net_edges$weight)*0.7),]
  
  gene_links<-table(c(net_edges$from,net_edges$to))
  filt_genes<-names(gene_links)[which(gene_links > links)] #select genes with link > links
  net_edges<-subset(net_edges,from %in% filt_genes & to %in% filt_genes) 
  
  gene_links<-table(c(net_edges$from,net_edges$to))
  filt_genes<-names(gene_links)[which(gene_links >= 2)] #remove isolated genes (leaves)
  net_edges<-subset(net_edges,from %in% filt_genes & to %in% filt_genes)
  
  net_edges$group_to<-"PP"
  
  # Vertices df ----
  gene_vertices<-data.frame(label=filt_genes)
  if(nrow(gene_vertices) != 0){
    rownames(gene_vertices)<-gene_vertices$label
    
    gene_vertices$Community<-(i_comms_df[[comp]])[as.character(gene_vertices$label),"comm_n"] #class=communities
    #Set size bigger for genes in a community
    gene_vertices$n_size<-1.5 #column that define the size of the dots
    gene_vertices[which(is.na(gene_vertices$Community)),"n_size"]<-0.75
    
    gene_vertices$type<-"A" # can be used for shape
    #gene_vertices[which(rownames(gene_vertices)%in%sma_genes),"type"]<-"B"
    
    nc<-ggplot(gene_vertices, aes(x=Community, fill=Community)) +
        geom_bar(alpha=0.9, colour="white", width=0.4)+
        scale_x_discrete(drop=FALSE)+
        scale_fill_manual(values=colour_vector[[comp]], drop=FALSE)+
        theme_bw(base_size = bs) +
        labs(x = "Community", y = "# genes") +
        theme(legend.position = "none", panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank(), panel.grid.minor.y = element_blank())+
        ggtitle("Number of genes in each community", subtitle = paste0("Communities of condition: ",comp))
    
    cat("##### ",comp," \n")
    print(nc)
    cat(" \n \n")
    ggsave(paste0(dirOutput_net,comp,"_community_number_gene_filtered.pdf"), nc, device=cairo_pdf, width = 4, height = 4, units = c("in"))
    
    # plot with DiagrammeR----
    net_net <- suppressMessages(fortify(as.edgedf(as.data.frame(net_edges)), gene_vertices))
    uniqueN(net_net$from_id)
    
    #BUILD NETWORK
    node_list_from<-as.data.table(unique(net_net[,c("from_id")]))
    node_list_to<-as.data.table(unique(net_net[,c("to_id")]))
    colnames(node_list_from)<-"node"
    colnames(node_list_to)<-"node"
    node_list<-as.data.table(rbind((node_list_from),(node_list_to)))
    node_list<-node_list[!duplicated(node_list$node)]
    node_list[,id:=seq(1, nrow(node_list))]
    node_list<-merge(node_list,net_net[,c("from_id","Community","n_size")],by.x="node",by.y="from_id")
    
    node_list<-unique(node_list)
    
    colnames(node_list)<-c("name","id","Community","size")
    
    edge_list<-net_net[,c("from_id","to_id","weight")]
    colnames(edge_list)<-c("from","to","weight")
    edge_list<-as.data.table(edge_list)
    edge_list<-na.omit(edge_list)
    g <- graph_from_data_frame(edge_list, directed=TRUE, vertices=node_list)
    
    name_list=vector()
    name_list<-foreach (l = c("fr","kk"), .packages = c("ggraph"), .combine = 'c') %dopar% {
      tryCatch({
        l_list<-ggraph(g,layout=l) +
          geom_edge_link2(aes(edge_width=(0.05+abs(weight-scr_thr)/400)),edge_colour = "grey50",alpha=0.2) +
          # ggtitle(paste0("Network of ",comp),subtitle = paste0("Layout: ",l))+
          scale_edge_width(range = c(0.1,1)) +
          geom_node_point(aes(fill = Community), shape = 21,size=1.5,pch='.') +
          geom_node_text(aes(label = name,size=5,color=Community),family = bf, repel = TRUE) +
          scale_fill_manual(values = colour_vector[[comp]]) +
          # scale_edge_width_continuous(name=c("Weigth","Size","Community")) +
          scale_color_brewer(palette ="Dark2") +
          theme_bw(base_size = bs, base_family = bf) +
          theme(legend.position = "bottom", panel.grid = element_blank(), 
                panel.border = element_blank(), panel.background = element_blank(),
                axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()) +
          guides(text=F)
        l_list$labels$edge_width<-"Weight"
        ggsave(paste0(dirOutput_net,comp,"_",l,"_network.pdf"), l_list, device=cairo_pdf, width = 10, height = 6, units = c("in"))
        # 
        print(paste0(dirOutput_net,comp,"_",l,"_network.pdf"))
        # name_list<-append(name_list,paste0(dirOutput_net,comp,"_",l,"_network.pdf"))
      }, error=function(cond){print("Error: Error occur during the plot of network.\n")})
    }
    tryCatch({
      pdf_combine(input = name_list, output = paste0(dirOutput_net,comp,"_network.pdf"))
      unlink(name_list)
    }, error=function(cond){print("Error: Not possible combine PDFs network in single file.\n")})
    
    cat(" \n \n")
  }
}
stopCluster(cluster_ext)

cat("\n \n")
message("Networking with STRING DONE")
cat("### 5.2. The networks can be found in the pdfs inside *network* directory")
cat("\n")
cat("The network are not reported in this report since for each comparison are generated 2 figures since we use 2 layouts for the nodes. For each comparison a 2-page pdf file is created. \n")

message("Cleaning...")
```


```{r save_df, echo=F, message=F, warning=F}
save(c_anno, 
     dat_gene, 
     dat_pep, 
     psm_peptide_table, 
     deps_l_df, 
     deps_pep_l_df, 
     expr_avgse_df, 
     expr_avgse_pep_df, 
     formule_contrast, 
     colour_vec,
     file = paste0(dirOutput,"database_env_R.RData"))
```